### 2017-02-01 states/resistance/report/state-report.Rmd
state report

---
title: 'Consumption and resistance across states'
---

```{r echo=FALSE, message=FALSE}
res = read_tsv('../model-results-p005.tsv')
plot_dat = read_tsv('../plotting-dat.tsv')

plot_f = function(this_bug, this_resistance_drug, this_explanatory, this_response) {
  this_res = res %>%
    filter(bug==this_bug, resistance_drug==this_resistance_drug, explanatory==this_explanatory, response==this_response)

  if (dim(this_res)[1] != 1) stop('wrong number of rows')

  z = this_res$correlation %>% head
  p.value = this_res$p.value %>% head

  plot_dat %>%
    filter(bug==this_bug, resistance_drug==this_resistance_drug) %>%
    ggplot(aes_string(this_explanatory, this_response)) +
    geom_point() +
    geom_smooth(method='lm', se=FALSE) +
    theme_minimal() +
    xlab(this_explanatory) +
    ylab(this_response) +
    ggtitle(sprintf('%s susceptibility to %s', this_bug, this_resistance_drug))
}
```

# Methodology

1. I determined which bug/drug combinations were coherent at the state level (i.e., separating antibiograms into their states significantly decreased variance below what would be expected by chance)
2. For each *consumption* drug, I asserted a matched *resistance* drug
3. I correlated metrics of consumption (of the consumption drug) against metrics of resistance (against the resistance drug) across states.

## Metrics

The consumption metrics were:

- *DID*: Defined daily doses per 1,000 inhabitants per day. In fact, I don't have proper DDDs, so these are the "days supply" supplied in the Medicare claims data. This is what Goossens *et al.* uses.
- *Claims per 1,000 people*: Rather than days supply, this is just the raw number of claims. It also has a different denominator. This is what the Hicks papers use.

In general, PKPY will be larger than DID. If all people get $d$ days of drug per prescription, the relationship is:

$$
\mathrm{PKPY} = \frac{365}{d} \times \mathrm{DID}
$$

The resistance metrics were:

- *Mean* susceptibility
- *Median* susceptibility
- *Maximum* susceptibility
- *Minimum* susceptibility

## Correlations

I used Spearman correlations and examined only those bug/drug combinations that produced $p < 0.05$. I manually examined the results and show some representative ones here.

## Consumption and resistance drugs

Trimethoprim-sulfamethoxazole, ciprofloxacin, levofloxacin, ceftriaxone, nitrofurantoin, and penicillin were present in both the consumption and resistance data; they are each their own pair. Consumption of amoxicillin was matched with resistance to ampicillin; doxycycline with tetracycline/doxycycline, and azithromycin with erythromycin.

## Organisms

I examined these organisms:

Gram negatives:
- Escherichia coli
- Klebsiella pneumoniae
- Proteus mirabilis
- Enterobacter cloacae
- Klebisella oxytoca

Gram positives:
- Coagulase-negative staphylococci
- Methicillin-resistant S. aureus (MRSA)
- Methicillin-susceptible S. aureus (MSSA)
- Streptococcus pneumoniae (non-meningitis)
- Staphylococcus aureus

# Results

## Strong results

I call these strong results because:

- Their p-values were small.
- The correlations went in the right direction (i.e., increased consumption was associated with decreased susceptibility).
- The same bug/drug/drug combinations had significant correlations for multiple metrics (e.g., *E. coli* susceptibility to cipro anticorrelated with cipro consumption regardless of whether consumption was measured with claims or days and whether the mean or median susceptibility was used).

In general, the coefficients for these correlations were in the range 0.4--0.6.

N.B.: The plots have a linear best fit line added for clarity, but the correlation was measured with Spearman (i.e., rank ordered).

```{r echo=FALSE}
plot_f('Escherichia coli', 'ciprofloxacin', 'claims_per_1k_ppl', 'mean_sus')
plot_f('Escherichia coli', 'levofloxacin', 'claims_per_1k_ppl', 'mean_sus')
plot_f('Methicillin-susceptible S. aureus (MSSA)', 'ciprofloxacin', 'claims_per_1k_ppl', 'mean_sus')
plot_f('Methicillin-susceptible S. aureus (MSSA)', 'erythromycin', 'claims_per_1k_ppl', 'mean_sus')
plot_f('Streptococcus pneumoniae (non-meningitis)', 'erythromycin', 'claims_per_1k_ppl', 'mean_sus')
```

## Dubious Results

I call these results "dubious" because, in general:

- The p-values were not as strong.
- The correlations were sometimes in the wrong direction.
- The susceptibility metrics were weird (e.g., consumption correlations with the minimum susceptibility reported in a state).
- The correlation was significant for one consumption metric but not the other.

```{r echo=FALSE}
plot_f('Coagulase-negative staphylococci', 'levofloxacin', 'claims_per_1k_ppl', 'min_sus') # why only min sus?
plot_f('Enterobacter cloacae', 'ceftriaxone', 'did', 'mean_sus') # why only DID?
plot_f('Escherichia coli', 'ampicillin', 'did', 'median_sus') # why only DID?
plot_f('Escherichia coli', 'trimethoprim-sulfamethoxazole', 'claims_per_1k_ppl', 'mean_sus') # why not claims per 1k ppl?
plot_f('Klebsiella pneumoniae', 'ceftriaxone', 'did', 'mean_sus') # why only DID?
plot_f('Methicillin-resistant S. aureus (MRSA)', 'levofloxacin', 'claims_per_1k_ppl', 'max_sus') # why only max sus?
plot_f('Methicillin-resistant S. aureus (MRSA)', 'trimethoprim-sulfamethoxazole', 'did', 'mean_sus') # why only DID?
plot_f('Proteus mirabilis', 'levofloxacin', 'claims_per_1k_ppl', 'min_sus') # why only min sus?
plot_f('Streptococcus pneumoniae (non-meningitis)', 'trimethoprim-sulfamethoxazole', 'claims_per_1k_ppl', 'mean_sus') # why only some?
```

# Discussion

## Comparison to Goossens

Goossens *et al.* found significant correlations for:

- Consumption of macrolides with *S. pneumoniae* susceptibility to erythromycin
- Consumption of penicillins and cephalosporins with *S. pneumoniae* susceptibility to penicillin
- Consumption of macrolides and lincosamids with *S. pyogenes* susceptibility to erythromycin
- Consumption of quinolones with *E. coli* susceptibility to ciprofloxacin
- Consumption of co-trimoxazole with *E. coli* susceptibility to co-trimoxazole

The Spearman correlation coefficients range from 0.68 to 0.83, which are higher than our results.

The Goossens *et al.* graph shows a best-fit line of penicillin consumption against *S. pneumoniae* non-susceptibility to penicillin: 5 DID corresponds to 0% non-susceptible, while 16 DID corresponds to 45% non-susceptible. The translates to a 4 percentage point change in susceptibility per 1 DID.

I can't immediately compare this to the US data (because *S. pneumoniae* resistance to penicillin was not coherent at the state level).

However, *E. coli* resistance to ciprofloxacin is interesting: 85% susceptibility at 50 PKPY and 75% at 200, that is, 1 percentage point in susceptibility per 15 PKPY. That's a very weak response!