### 2017-02-03 states/resistance/report2/state-report.Rmd
second report

---
title: "Consumption and resistance across states"
output:
  html_document:
    fig_height: 4
  pdf_document:
    fig_height: 3
---

```{r setup}
library(broom)
library(knitr)
library(pander)

coherence = read_tsv('../../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
  rename(resistance_drug=drug) %>%
  mutate(resistance_drug=tolower(resistance_drug)) %>%
  select(bug, resistance_drug, level, n_groups)

gram_negs = c('Escherichia coli', 'Klebsiella pneumoniae', 'Proteus mirabilis',
              'Enterobacter cloacae', 'Klebisella oxytoca')
gram_poss = c('Coagulase-negative staphylococci', 'Methicillin-resistant S. aureus (MRSA)',
              'Methicillin-susceptible S. aureus (MSSA)', 'Streptococcus pneumoniae (non-meningitis)',
              'Staphylococcus aureus')
organisms = c(gram_negs, gram_poss)

consumption_drugs=c('trimethoprim-sulfamethoxazole', 'ciprofloxacin', 'levofloxacin',
                    'ceftriaxone', 'nitrofurantoin', 'amoxicillin', 'penicillin',
                    'doxycycline', 'azithromycin')
resistance_drugs=c('trimethoprim-sulfamethoxazole', 'ciprofloxacin', 'levofloxacin',
                   'ceftriaxone', 'nitrofurantoin', 'ampicillin', 'penicillin',
                   'tetracycline/doxycycline', 'erythromycin')

consumption_resistance = data_frame(consumption_drug=consumption_drugs,
                                    resistance_drug=resistance_drugs)

# get all bug/drug combos (i.e., all bug/cons drug combos with appropriate res drugs)
bug_drug = expand.grid(bug=organisms,
                       consumption_drug=consumption_drugs,
                       stringsAsFactors=FALSE) %>%
  left_join(consumption_resistance, by='consumption_drug') %>%
  left_join(coherence, by=c('resistance_drug', 'bug')) %>%
  mutate(state_coherent=level %in% c('state', 'hrr', 'zipcode'))
```

```{r load_data}
abg = read_tsv('../../../../../antibiogram/data/abg.tsv') %>%
  mutate(drug=tolower(drug)) %>%
  rename(resistance_drug=drug)

known_states = abg %$% state %>% unique %>% sort

state_bene = read_tsv('../../bene-consumption/state-bene-2011.tsv') %>%
  filter(state %in% known_states)
state_cons = read_tsv('../../bene-consumption/state-consumption-2011.tsv') %>%
  filter(state %in% known_states) %>%
  rename(consumption_drug=drug)

sus = bug_drug %>%
  left_join(abg, by=c('bug', 'resistance_drug')) %>%
  group_by(bug, resistance_drug, state) %>%
  summarize(n_abg=n(),
            n_isolates=sum(n_isolates, na.rm=TRUE),
            mean_sus=mean(percent_susceptible, na.rm=TRUE))
```

```{r run_models}
safe_spearman = function(df, expl) {
  if (dim(df)[1] < 3L) {
    data_frame(estimate=NA, statistic=NA, p.value=NA, method=NA, alternative=NA)
  } else {
    cor.test(df[['mean_sus']], df[[expl]], method='spearman') %>% tidy
  }
}

models = function(df) {
  out = data_frame()
  for (expl in c('claims_per_1k_ppl', 'did')) {
    new_rows = safe_spearman(df, expl) %>% 
      mutate(explanatory=expl)

    out = bind_rows(out, new_rows)
  }
  out
}

dat = sus %>%
  left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
  left_join(state_cons, by=c('state', 'consumption_drug')) %>%
  group_by(bug, resistance_drug)

res = dat %>%
  do(models(.))
```

```{r echo=FALSE, message=FALSE}

plot_f = function(this_bug, this_resistance_drug, this_explanatory, this_response) {
  this_res = res %>%
    filter(bug==this_bug, resistance_drug==this_resistance_drug, explanatory==this_explanatory, response==this_response)

  if (dim(this_res)[1] != 1) stop('wrong number of rows')

  z = this_res$correlation %>% head
  p.value = this_res$p.value %>% head

  plot_dat %>%
    filter(bug==this_bug, resistance_drug==this_resistance_drug) %>%
    ggplot(aes_string(this_explanatory, this_response)) +
    geom_point() +
    geom_smooth(method='lm', se=FALSE) +
    theme_minimal() +
    xlab(this_explanatory) +
    ylab(this_response) +
    ggtitle(sprintf('%s susceptibility to %s', this_bug, this_resistance_drug))
}
```

```{r}
res %>%
  head %>%
  select(bug, resistance_drug, explanatory, correlation=estimate, p.value) %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  kable
```