### 2017-02-15 temporal/seasonal-traces/seasonal-trace.Rmd
seasonal

---
title: "Seasonal patterns for azithromycin across states"
author: "Scott Olesen"
date: "2/15/2017"
output: html_document
---

# Motivation

Azithromycin has a seasonal pattern in prescriptions. Does that pattern vary by geography? I used the Medicare data.

```{r load_data, echo=FALSE, message=FALSE}
dat2011 = read_tsv('../temporal-abx-2011.tsv')

full_dat = lapply(2011:2014, function(x) { read_tsv(sprintf('../temporal-abx-%i.tsv', x)) %>% mutate(year=x) }) %>%
  bind_rows()
```

# Data

By eye, it looks like the peak in consumption is similar across states (it occurs just after New Year's), and the minimum occurs at around the same time (weeks 27-36, which are the months of July and August).

```{r plot_azithro_two_states, echo=FALSE}
tick_labels = seq(1, 52) %>% { if_else(. %% 2 == 1, as.character(.), '') }
dat2011 %>%
  filter(abx == 'azithromycin', state %in% c('Florida', 'New York', 'Texas', 'California', 'Michigan'), week < 53.0) %>%
  ggplot(aes(x=factor(week), y=n_rx, group=state, color=state)) +
  geom_point() + geom_line() +
  xlab('week of year 2011') + ylab('# of claims') +
  scale_x_discrete(labels=tick_labels) +
  theme_minimal() +
  ggtitle('Azithromycin claims across states in 2011')
```

The pattern within a state looks fairly consistent across years

```{r plot_azithro_years, echo=FALSE}
full_dat %>%
  filter(abx == 'azithromycin', state == 'Florida', week < 53.0) %>%
  mutate(year=factor(year)) %>%
  ggplot(aes(x=factor(week), y=n_rx, group=year, color=year)) +
  geom_point() + geom_line() +
  xlab('week of year') + ylab('# of claims') +
  scale_x_discrete(labels=tick_labels) +
  theme_minimal() +
  ggtitle('Azithromycin claims in Florida across years')
```

I wrote a function to find the time until the number of prescriptions falls by half. Specifically, I:

- fit a smoothed spline (loess) to timepoints between weeks a start week and end week
- find the mean value between the predicted (i.e., the spline's values) for between the start and end weeks
- find the time at which the spline hits that value

```{r midpoint_function}
start_week = 5 # after the fast dropoff from New Years
end_week = 27 # just before the July/August lull

midpoint_time = function(n_rx, week) {
  f = loess(n_rx ~ week)
  midval = mean(c(predict(f, start_week), predict(f, end_week)))
  uniroot(function(x) predict(f, x) - midval, c(start_week, end_week))$root
}
```

It looks like there is some relationship between geography and midpoint week: in the northernmost states, consumption falls to half about 5 weeks after the southernmost states. The values are super-beautiful, but the metric I used to measure them is pretty janky.

```{r show_midpoints, echo=FALSE}
library(knitr)

full_dat %>%
  filter(abx == 'azithromycin', state %in% c('Florida', 'New York', 'Texas', 'California', 'Michigan', 'North Dakota', 'Mississippi'), week <= end_week) %>%
  group_by(year, state) %>%
  summarize(midpoint_time=midpoint_time(n_rx, week)) %>%
  mutate(midpoint_time=round(midpoint_time, 1)) %>%
  spread(year, midpoint_time) %>%
  arrange(desc(`2011`)) %>%
  kable
```