### 2017-02-16 verification/all-pde-days/all-pde-days.Rmd
pde by day

---
title: "PDE special days"
author: "Scott Olesen"
date: "2/16/2017"
output: html_document
---

The PDE data has funny bumps and wiggles. I think these are due to things like the start of the month or holidays.

# Data

Using SAS on the cluster, I collected the counts for each unique date among all the PDEs.

# Results

```{r load_data, message=FALSE, echo=FALSE}
library(lubridate)
library(knitr)
dow_levels = c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday')
d11 = read_tsv('pde_date_counts_2011.tsv') %>%
  mutate(date=dmy(date)) %>%
  mutate(week_n=week(date)) %>%
  mutate(day_of_week=date %>% weekdays %>% factor(levels=dow_levels, ordered=TRUE)) %>%
  group_by(day_of_week) %>%
  mutate(day_of_week_z_score=as.vector(scale(count))) %>%
  ungroup()
```

There is a clear pattern by day:
```{r plot_days}
d11 %>%
  ggplot(aes(x=date, y=count)) +
  geom_line() + geom_point() +
  xlab('day of the year 2011') + ylab('PDEs on that day') +
  theme_minimal()
```

It looks like there are a few things going on in that day:

- a fast, weekly rhythm
- a slower, monthly rhythm
- a few weirdly low points

# Patterns within a week

It looks like there is a pattern within each week:
```{r plot_weekday, echo=FALSE}
d11 %>%
  ggplot(aes(x=day_of_week, y=count)) +
  geom_boxplot() +
  xlab('') + ylab('PDEs on that day') +
  theme_minimal()
```

Some of these days fall outside the normal range. For example, there are three or four Mondays and one Thursday that fall far below the normal range.
```{r outliers_plot, echo=FALSE}
d11 %>%
  ggplot(aes(x=day_of_week, y=day_of_week_z_score)) +
  geom_boxplot() +
  xlab('day of week') + ylab('z score (within the week)') +
  theme_minimal()
```

Which days are the ones that have large z-scores?
```{r outliers}
d11 %>%
  select(day_of_week, date, day_of_week_z_score) %>%
  filter(abs(day_of_week_z_score) >= 2) %>%
  arrange(day_of_week) %>%
  kable
```

The dates with abnormally low numbers were:

- April 24 (Easter Sunday)
- May 30 (Memorial Day)
- July 4
- September 5 (Labor Day)
- November 24 (Thanksgiving)
- December 24 (Christmas Eve)
- December 25 (Christmas)

The dates with abnormally high numbers of PDEs were all the first of the month or the first work day after a holiday:

- January 2 (the Sunday after New Year's)
- January 3 (the first Monday after New Year's)
- April 1 (Friday)
- May 1 (Sunday)
- May 31 (the Tuesday after Memorial Day Monday)
- June 1 (Wednesday)
- July 1 (Friday)
- July 5 (the Tuesday after July 4 Monday)
- September 1 (Thursday)
- September 6 (the Tuesday after Labor Day Monday)
- October 1 (Saturday)
- December 1 (Thursday)

The first-of-months that are missing from this list are all Mondays and Tuesdays (i.e., otherwise busy days):

- February 1 (Tuesday)
- March 1 (Tuesday)
- August 1 (Monday)
- November 1 (Tuesday)

# Patterns across weeks

There is variation week-to-week:
```{r plot_week}
d11 %>%
  group_by(week_n) %>%
  summarize(count=sum(count)) %>%
  ggplot(aes(x=week_n, y=count)) +
  geom_point() + geom_line() +
  xlab('week of year 2011') + ylab('count') +
  theme_minimal()
```

The most obvious features are the weird last week and an apparently monthly pattern.

The last week is weird because it has only one day in it:
```{r last_week}
d11 %>% filter(week_n==53)
```

# Patterns across months

What about the monthly pattern? Which are the weeks that 
```{r exceptional_weeks}
summarize_week = function(x) {
  data_frame(start_day=min(x$date),
             end_day=max(x$date),
             count=sum(x$count))
}

d11 %>%
  group_by(week_n) %>%
  do(summarize_week(.)) %>%
  ungroup() %>%
  mutate(week_z_score=as.vector(scale(count))) %>%
  filter(abs(week_z_score) > 0.25) %>%
  arrange(week_z_score) %>%
  kable
```

The three lowest weeks are the weird last day, the week of Thanksgiving, and the week of Christmas. The highest weeks are:

- first week of December
- first week of January
- first week of March
- first week of October

and so forth.