### 2017-02-27 inequality/consumption-inequality.Rmd
fixed error

---
title: "Inequality in consumption"
author: "Scott Olesen"
date: "2/22/2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/',
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

# Background

What is the inequality in consumption for antibiotics? How does it depend on region or the antibiotic used?

# Methods

I define the Juran inequality index $J$: for a random variable $X$ with a domain $\{0, 1\}$ and a cumulative distribution function $F_X$, $J(X)$ is that value such that $F(J(X)) = 1 - J(X)$. For example, if the top 20% of beneficiaries make 80% of the claims, then $J = 0.8$.

For each group under consideration, I considered 5 different metrics:

- the fraction of beneficiaries with at least one claim
- the Juran index among all beneficiaries
- the Juran index among beneficiaries with at least one claim
- the Gini index among all beneficiaries
- the Gini index among beneficiaries with at least one claim

```{r setup}
library(ineq)
is_nonzero = function(x) x > 0
nonzero = function(x) x[is_nonzero(x)]
fraction_nonzero = function(x) sum(is_nonzero(x)) / length(x)
gini = function(x) ineq(x, type='Gini')

normalize = function(x) x / max(x)
juran = function(x) {
  # better to do with percent_rank and cume_dist
  fraction_consumed = cumsum(sort(x, decreasing=TRUE)) %>% normalize
  fraction_users = 1:length(x) %>% normalize
  fraction_nonzero(fraction_consumed + fraction_users - 1)
}

add_zeros = function(x, n) {
  if (n < length(x)) {
    stop(sprintf('cannot extend x of length %i to %i', length(x), n))
  }

  c(x, rep(0, n - length(x)))
}

summarize_claims = function(x) {
  summarize(x, n_claims=sum(n_claims), n_days=sum(days_supply)) %>%
    mutate(n_days=if_else(is.na(n_days), 0L, n_days))
}

summarize_inequality = function(x, ...) {
  do(x, inequalities(., ...))
}

inequalities = function(x, what, denom=NULL) {
  z = x[[what]]
  if (!is.null(denom)) z = add_zeros(z, denom)

  data_frame(fraction_nonzero=fraction_nonzero(z),
             juran=juran(z),
             juran_nonzero=juran(nonzero(z)),
             gini=gini(z),
             gini_nonzero=gini(nonzero(z)))
}

top_antibiotics = c('azithromycin', 'ciprofloxacin', 'levofloxacin', 'amoxicillin', 'cephalexin', 'trimethoprim/sulfamethoxazole', 'nitrofurantoin', 'doxycycline')
```

# Results

```{r load_data}
usage = read_tsv('../bene_usage_2011.tsv')
```

Overall, about half of beneficiaries took at least one antibiotic. Overall, 20% of beneficiaries claimed 80% of the antibiotic days' supply.

```{r inequality_overall}
# put in both metrics of consumption
# run it through summarize_inequality
# then group by region or by drug also
usage %>%
  group_by(bene_id) %>%
  summarize_claims() %>%
  gather('consumption_metric', 'amount', n_claims, n_days) %>%
  group_by(consumption_metric) %>%
  summarize_inequality('amount') %>%
  kable
```

```{r inequality_hist}
usage %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims)) %>%
  ggplot(aes(x=n_claims)) +
  geom_histogram() +
  scale_y_log10() +
  xlab('number of claims') + ylab('number of beneficiaries') +
  theme_minimal()
```

Breaking down these results by region shows that the South, which consumed the most antibiotic per capita, has a slightly more even usage (Juran index of 79% versus 80% in other regions). I'm not sure that's a meaningful difference.

```{r inequality_by_region}
usage %>%
  group_by(region, bene_id) %>%
  summarize_claims() %>%
  gather('consumption_metric', 'amount', n_claims, n_days) %>%
  group_by(region, consumption_metric) %>%
  summarize_inequality('amount') %>%
  kable
```

Among the most popular drugs, unevenness varies only slightly. Among users, azithromycin consumption has an approximately 60% Juran index, and ciprofloxacin has 65%, while TMP-SMX and vancomycin have 70%.

```{r inequality_by_drug}
n_bene = usage$bene_id %>% unique %>% length
usage %>%
  group_by(antibiotic, bene_id) %>%
  summarize_claims() %>%
  gather('consumption_metric', 'amount', n_claims, n_days) %>%
  group_by(antibiotic, consumption_metric) %>%
  summarize_inequality('amount', n_bene) %>%
  filter(antibiotic %in% top_antibiotics) %>%
  kable
```

## Inequality by state
```{r compute_inequality_by_state}
inequality_by_state = usage %>%
  group_by(state, bene_id) %>%
  summarize_claims() %>%
  gather('consumption_metric', 'amount', n_claims, n_days) %>%
  group_by(state, consumption_metric) %>%
  summarize(n_bene=n(),
             fraction_nonzero=fraction_nonzero(amount),
             juran=juran(amount),
             juran_nonzero=juran(nonzero(amount)),
             gini=gini(amount),
             gini_nonzero=gini(nonzero(amount))) %>%
  ungroup()
```

The distribution of inequalities seems small: all states have (non-zero) Juran
indices that fall between about 0.7 and 0.75. That being said, there does seem
to be a trend by region: the fraction nonzero is highest in the South?

```{r show_inequality_by_state}
inequality_by_state %>%
  filter(consumption_metric=='n_days') %>%
  ggplot(aes(x=juran_nonzero)) + geom_histogram() +
  theme_minimal()

inequality_by_state %>%
  filter(fraction_nonzero > 0) %>%
  filter(consumption_metric=='n_days') %>%
  select(-consumption_metric) %>%
  arrange(state) %>%
  kable
```

## Characteristics by consumption

I broke down the beneficiaries based on their consumption: those who had no antibiotic claims were in one "tile", and I separated the rest into $n$-tiles.

```{r chars_by_consumption}
ntile_zero = function(x, n) {
  out = x
  out[x > 0] = ntile(out[x > 0], n)
  out
}

bounds_group = function(x, upper_limits) {
  upper_limits = as.integer(upper_limits)
  levels = lapply(seq(from=1, to=length(upper_limits)), function(i) {
    if (i == 1) {
      sprintf('<= %i', upper_limits[i])
    } else {
      sprintf('> %i and <= %i', upper_limits[i - 1], upper_limits[i])
    }
  })
  top_level = sprintf('> %i', tail(upper_limits, 1))
  formula_list = mapply(function(str, lim) as.formula(sprintf('x <= %i ~ "%s"', lim, str)),
                        levels, upper_limits) %>%
    c(as.formula(sprintf('TRUE ~ "%s"', top_level)))

  levels = c(levels, top_level)

  do.call(case_when, formula_list) %>%
    factor(levels=levels, ordered=TRUE)
}

usage %>%
  group_by(age, sex, race, chronic, bene_id) %>%
  summarize(n_claims=sum(n_claims),
            n_days=sum(n_days)) %>%
  ungroup() %>%
  mutate(claims_group=bounds_group(n_claims, c(0, 1, 2, 5, 10, 20, 50, 100)),
         is_female=if_else(sex=='female', 1, 0),
         is_white=if_else(race=='white', 1, 0)) %>%
  group_by(claims_group) %>%
  summarize(n_beneficiaries=n(),
            mean_age=mean(age),
            percent_female=mean(is_female)*100,
            percent_white=mean(is_white)*100,
            mean_n_chronic_cond=mean(chronic),
            mean_n_claims=mean(n_claims)) %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  kable
```

# Discussion

I used only the 2011 data in this analysis. I intend to extend the analysis to other years. Data from other years could help confirm any of the (relatively small) differences in inequality in consumption.

Surprisingly, the characteristics of the beneficiaries are fairly constant by consumption $n$-tile among those who took antibiotics. Those who took no antibiotics where more male and less white.