### 2017-02-28 states/resistance/select_plots.Rmd
select plots for presentation

---
title: "Select consumption vs resistance plots"
bibliography: refs.bib
output:
  html_document:
    fig_height: 4
  pdf_document:
    fig_height: 3
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r load_data}
# Antibiogram data, with individual records
abg = read_tsv('../../../../antibiogram/data/abg.tsv') %>%
  mutate(drug=tolower(drug)) %>%
  rename(resistance_drug=drug) %>%
  select(bug, resistance_drug, state, percent_susceptible)

# State-level consumption data
state_consumption = read_tsv('../bene-consumption/state_usage_2011.tsv') %>%
  rename(consumption_drug=antibiotic) %>%
  select(state, consumption_drug, cpkp, did)
```

E. coli against cipro resistance and consumption. First plot is against state means; second includes all data points.

```{r ecoli_cipro}
cons = state_consumption %>%
  filter(consumption_drug=='ciprofloxacin')

res = abg %>%
  filter(resistance_drug=='ciprofloxacin', bug=='Escherichia coli') %>%
  mutate(pct_nonsus=100-percent_susceptible)

res_sum = res %>%
  group_by(state) %>%
  summarize(mean_pct_nonsus=mean(pct_nonsus))

full_join(cons, res_sum, by='state') %>%
  filter(!is.na(mean_pct_nonsus), !is.na(cpkp)) %>%
  ggplot(aes(x=cpkp, y=mean_pct_nonsus)) +
  geom_point() +
  geom_smooth(method='glm', se=FALSE) +
  ylim(c(0, 50)) +
  xlab('ciprofloxacin claims per 1,000 people per year') + ylab('mean percent cirpo. non-susceptible E. coli') +
  theme_minimal()
```

```{r ecoli_cipro_all}
dat_all = bind_rows(
  res %>% mutate(summary=FALSE),
  res_sum %>% rename(pct_nonsus=mean_pct_nonsus) %>% mutate(summary=TRUE)
) %>% 
  full_join(cons, by='state') %>% 
  filter(!is.na(pct_nonsus), !is.na(cpkp))
  ggplot(dat_all %>% filter(summary), aes(x=cpkp, y=pct_nonsus)) +
  geom_point() +
  geom_point(dat=dat_all %>% filter(!summary), shape=1, size=0.5) +
  geom_smooth(dat=dat_all %>% filter(!summary), method='glm', se=FALSE, col='gray') +
  geom_smooth(method='glm', se=FALSE) +
  ylim(c(0, 75)) +
  xlab('ciprofloxacin claims per 1,000 people per year') + ylab('percent cipro. non-susceptible E. coli') +
  theme_minimal()
```

S. pneumoniae against azithromycin consumption and erythromycin resistance.

```{r strep_azi}
cons = state_consumption %>%
  filter(consumption_drug=='azithromycin')

res = abg %>%
  filter(resistance_drug=='erythromycin', bug=='Streptococcus pneumoniae (non-meningitis)') %>%
  mutate(pct_nonsus=100-percent_susceptible)

res_sum = res %>%
  group_by(state) %>%
  summarize(mean_pct_nonsus=mean(pct_nonsus))

full_join(cons, res_sum, by='state') %>%
  filter(!is.na(mean_pct_nonsus), !is.na(cpkp)) %>%
  ggplot(aes(x=cpkp, y=mean_pct_nonsus)) +
  geom_point() +
  geom_smooth(method='glm', se=FALSE) +
  ylim(c(0, 75)) +
  xlab('azithromycin claims per 1,000 people per year') + ylab('mean percent azith. non-susceptible S. pneumoniae') +
  theme_minimal()
```

```{r strep_azi_all}
dat_all = bind_rows(
  res %>% mutate(summary=FALSE),
  res_sum %>% rename(pct_nonsus=mean_pct_nonsus) %>% mutate(summary=TRUE)
) %>% 
  full_join(cons, by='state') %>% 
  filter(!is.na(pct_nonsus), !is.na(cpkp))
  ggplot(dat_all %>% filter(summary), aes(x=cpkp, y=pct_nonsus)) +
  geom_point() +
  geom_point(dat=dat_all %>% filter(!summary), shape=1, size=0.5) +
  geom_smooth(dat=dat_all %>% filter(!summary), method='glm', se=FALSE, col='gray') +
  geom_smooth(method='glm', se=FALSE) +
  ylim(c(0, 75)) +
  xlab('azithromycin claims per 1,000 people per year') + ylab('percent eryth. non-susceptible S. pneumoniae') +
  theme_minimal()
```