### 2017-03-06 inequality-resistance/inequality-resistance.Rmd
beginning ineq res

---
title: ""
author: "Scott Olesen"
date: ""
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, 
                      echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE)
```

First, load the relevant data. I need the usage and antibiograms.

```{r load_groups}
resistance_groups = read_tsv('../states/resistance/resistance_groups.tsv')
consumption_groups = read_tsv('../states/resistance/consumption_groups.tsv')
```

```{r load_antibiogram}
organisms = c('Escherichia coli')

hrr = read_tsv('../../db/hrr/hrr_2011.tsv') %>%
  select(zipcode, hrr)

abg = read_tsv('../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
  mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
  left_join(hrr, by='zipcode') %>%
  right_join(resistance_groups, by='drug') %>%
  filter(bug %in% organisms) %>%
  select(bug, drug_group, state, hrr, percent_nonsusceptible, n_isolates)

# Keep only those bug/drug combinations that are in at least 20 states
abg %<>%
  select(bug, drug_group, state) %>%
  distinct %>%
  count(bug, drug_group) %>% rename(n_states=n) %>%
  filter(n_states >= 20) %>%
  left_join(abg, by=c('bug', 'drug_group'))
```

```{r load_consumption}
usage = read_tsv('../bene_usage_2011.tsv') %>%
  rename(drug=antibiotic)

denom = usage %>% select(bene_id, state) %>% distinct

consumption = usage %>%
  left_join(consumption_groups, by='drug') %>%
  group_by(bene_id, drug_group) %>%
  summarize(n_claims=sum(n_claims)) %>%
  ungroup

consumption_of = function(drug_group0) {
  consumption %>%
    filter(drug_group==drug_group0) %>%
    select(bene_id, n_claims) %>%
    right_join(denom, by='bene_id') %>%
    mutate(n_claims=if_else(is.na(n_claims), 0L, n_claims))
}
```

```{r inequality_functions}
library(ineq)
is_nonzero = function(x) x > 0
nonzero = function(x) x[is_nonzero(x)]
fraction_nonzero = function(x) sum(is_nonzero(x)) / length(x)
gini = function(x) ineq(x, type='Gini')

normalize = function(x) x / max(x)
juran = function(x) {
  # better to do with percent_rank and cume_dist
  fraction_consumed = cumsum(sort(x, decreasing=TRUE)) %>% normalize
  fraction_users = 1:length(x) %>% normalize
  fraction_nonzero(fraction_consumed + fraction_users - 1)
}
```