### 2017-03-08 inequality-resistance/inequality-resistance.Rmd
many changes

---
title: ""
author: "Scott Olesen"
date: ""
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, 
                      echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE)
```

First, load the relevant data. I need the usage and antibiograms.

```{r load_groups}
resistance_groups = read_tsv('../states/resistance/resistance_groups.tsv')
consumption_groups = read_tsv('../states/resistance/consumption_groups.tsv')
drug_groups = resistance_groups$drug_group
```

```{r load_antibiogram}
organisms = c('Escherichia coli',
              'Proteus mirabilis', 'Enterobacter cloacae',
              'Enterococcus faecium',
              'Coagulase-negative staphylococci',
              'Methicillin-resistant S. aureus (MRSA)',
              'Methicillin-susceptible S. aureus (MSSA)',
              'Streptococcus pneumoniae (non-meningitis)',
              'Staphylococcus aureus')

hrr = read_tsv('../../db/hrr/hrr_2011.tsv') %>%
  select(zipcode, hrr)

abg = read_tsv('../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
  mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
  left_join(hrr, by='zipcode') %>%
  right_join(resistance_groups, by='drug') %>%
  filter(bug %in% organisms) %>%
  select(bug, drug_group, state, hrr, percent_nonsusceptible, n_isolates)

# Keep only those bug/drug combinations that are in at least 20 states
abg %<>%
  select(bug, drug_group, state) %>%
  distinct %>%
  count(bug, drug_group) %>% rename(n_states=n) %>%
  filter(n_states >= 20) %>%
  left_join(abg, by=c('bug', 'drug_group'))
```

```{r load_consumption}
usage = read_tsv('../bene_usage_2011.tsv') %>%
  rename(drug=antibiotic)

denom = usage %>% select(bene_id, state) %>% distinct

consumption = usage %>%
  left_join(consumption_groups, by='drug') %>%
  group_by(bene_id, drug_group) %>%
  summarize(n_claims=sum(n_claims)) %>%
  ungroup %>%
  complete(bene_id, drug_group, fill=list('n_claims'=0L)) %>%
  left_join(denom, by='bene_id')

consumption_of = function(drug_group0) {
  consumption %>%
    filter(drug_group==drug_group0) %>%
    select(bene_id, n_claims) %>%
    right_join(denom, by='bene_id') %>%
    mutate(n_claims=if_else(is.na(n_claims), 0L, n_claims))
}
```

```{r inequality_functions}
library(ineq)
is_nonzero = function(x) x > 0
nonzero = function(x) x[is_nonzero(x)]
fraction_nonzero = function(x) sum(is_nonzero(x)) / length(x)
gini = function(x) ineq(x, type='Gini')

normalize = function(x) x / max(x)
juran = function(x) {
  # better to do with percent_rank and cume_dist
  fraction_consumed = cumsum(sort(x, decreasing=TRUE)) %>% normalize
  fraction_users = 1:length(x) %>% normalize
  fraction_nonzero(fraction_consumed + fraction_users - 1)
}

klasma = function(x, threshold) {
  if (!between(threshold, 0.0, 1.0)) stop("bad threshold")
  data_frame(x=sort(x, decreasing=TRUE)) %>%
    mutate(csx=cumsum(x), cd=cume_dist(csx)) %$%
    sum(x[cd <= 0.1]) / sum(x)
}
```

```{r}
mean_abg = abg %>%
  group_by(bug, drug_group, state) %>%
  summarize(mpr=mean(percent_nonsusceptible),
            wmpr=weighted.mean(percent_nonsusceptible, sqrt(n_isolates)))

state_consumption = consumption %>%
  group_by(state, drug_group) %>%
  summarize(f_nz=fraction_nonzero(n_claims),
            pct_nz=100*f_nz,
            juran0=juran(n_claims),
            juran_nz=juran(nonzero(n_claims)),
            gini0=gini(n_claims),
            gini_nz=gini(nonzero(n_claims)),
            k10_nz=klasma(nonzero(n_claims), 0.1),
            n_bene=n(),
            n_claims=sum(n_claims),
            cpkp=n_claims/n_bene*1000,
            cpku=cpkp/f_nz)

```

```{r do_models}
models = function(x, independent) {
  univariate_formula = paste(independent, '~ cpkp') %>% formula
  univariate_results = lm(univariate_formula, data=x) %>%
    anova %>% tidy %>% mutate(model='univariate')
  multivariate_formula = paste(independent, '~ pct_nz + cpku') %>% formula
  multivariate_results = lm(multivariate_formula, data=x) %>%
    anova %>% tidy %>% mutate(model='multivariate')
  bind_rows(univariate_results, multivariate_results)
}

results = left_join(mean_abg, state_consumption, by=c('state', 'drug_group')) %>%
  group_by(bug, drug_group) %>%
  do(models(., 'mpr')) %>%
  select(bug, drug_group, model, term, estimate, p.value)

results2 = left_join(abg, state_consumption, by=c('state', 'drug_group')) %>%
  group_by(bug, drug_group) %>%
  do(models(., 'percent_nonsusceptible')) %>%
  select(bug, drug_group, model, term, estimate, p.value)
```

# Results

## Correlations between metrics

In general,

- CPKP correlates strongly with `f_nz` (makes sense, since most users take ~1 rx)
- Intensive metrics correlate with one another (`gini_nz` and `juran_nz`)
- Extensive metrics correlate (e.g., `juran` with `f_nz`)
- Intensive and extensive metrics correlation (`juran_nz` with `f_nz`)

Thus, in general, increased consumption goes with increased inequality.

## Klasma 10

Klasma 10 is the fraction of the drugs consumed by the top 10% of users. Consider bug/drug combos where `k10_nz` in the model `percent_susceptible` (or mean r) `~ pct_nz + k10_nz` is significant. Those are all either tmp-smx or tetracycline, and in each case, the estimate for `cpkp` in the univariate model matches the sign of `pct_nz` in the multivariate model, and `k10_nz` in the multivariate model is always negative.

## Using CPKU

I tried to use CPKU as an intensive measure: what's the average consump (CPK) among users (U)? I got similar results to when I was using other intensive metrics.

## f_nz vs. CPKP

In general, the fraction nonzero is a better predictor of resistance than CPKP, i.e., the extensive metric is better than the garbled intensive metric.