### 2017-03-13 abx-abx-correlations/drug-drug.Rmd
tetrachoric

---
title: "Drug-drug correlations"
author: "Scott Olesen"
date: "8 March 2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, 
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

# Background

Does taking one drug make you more likely to take another one? Or less likely?

# Methods

I use the 2011 data and tetrachoric correlations. I look at only the top 10 drugs (and I group amox and amox/clav under amox).

## Tetrachoric correlation

Consider two dichotomous variables $X$ and $Y$. Each joint observation ($x_i$, $y_i$) can be one of $(0, 0)$, $(0, 1)$, $(1, 0)$, or $(1, 1)$. The tetrachoric correlation finds a correlation $\rho$ by finding the "cutpoints" $tau_X$ and $tau_Y$ such that the bivariate standard normal distribution with correlation $\rho$ has "quadrants" with masses that best match the relative frequencies of the joint observations. For example, the "lower-left" quadrant has mass

$$
\int_{-\infty}^{\tau_Y} \int_{-\infty}^{\tau_X} \mathcal{N}(x, y; \rho) \,\mathrm{d}x \,\mathrm{d}y
$$

and should match the relative frequency of observations $(0, 0)$. The $\rho$ of the best-fitting bivariate normal distribution is the $\rho$ for those two variables. For more than two variables, a multivariate normal distribution is used.

This method can be extend to a *polychoric* method where the variables are ordinal (rather than dichotomous) by adding multiple cutpoints for the various categories.

```{r load_data}
usage = read_tsv('../bene_usage_2011.tsv') %>%
  select(bene_id, drug=antibiotic)

denom = usage %>% select(bene_id) %>% distinct
```

# Results

```{r set_drug_levels}
drugs = c(
  'ciprofloxacin',
  'trimethoprim/sulfamethoxazole',
  'nitrofurantoin',
  'cephalexin',
  'metronidazole',
  'levofloxacin',
  'azithromycin',
  'clindamycin',
  'amoxicillin', 'amoxicillin/clavulanate',
  'doxycycline'
)

drugs_short = c(
  'cpro', 'tpsx', 'ntro', 'cphx', 'mtro', 'levo', 'azth', 'clnd', 'amox',
  'amox/clav', 'doxy')
```

```{r do_correlation}
library(psych)
# convert all amox/clav entries into amox
res = usage %>%
  mutate(drug=if_else(
    drug %in% c('amoxicillin', 'amoxicillin/clavulanate'),
    'amoxicillin', drug)) %>%
  filter(drug %in% drugs) %>%
  distinct %>% # only care if had at least one claim
  right_join(denom, by='bene_id') %>% # keep benes with no claims
  mutate(has=TRUE) %>% # fill with TRUEs
  spread(drug, has, fill=FALSE) %>% # but with FALSE is they don't have
  select(-`<NA>`) %>% # drop the no-having column
  select(-bene_id) %>%
  tetrachoric
```

First, a table of the results.

```{r show_table}
res$rho %>%
  as.data.frame() %>%
  mutate(drug1=rownames(.)) %>%
  mutate_at(vars(-drug1), function(x) round(x, 2)) %>%
  select(drug1,
         amox=amoxicillin,
         azth=azithromycin,
         cphx=cephalexin,
         cipr=ciprofloxacin,
         clnd=clindamycin,
         doxy=doxycycline,
         levo=levofloxacin,
         mtro=metronidazole,
         ntro=nitrofurantoin,
         tpsx=`trimethoprim/sulfamethoxazole`) %>%
  kable
```

Then, a heatmap showing the same things. I picked the order of the rows/columns by hand.

```{r show_heatmap}
drug_factor = function(x) factor(x, levels=drugs, labels=drugs_short, ordered=TRUE)

res$rho %>%
  as.data.frame %>%
  mutate(drug1=rownames(.)) %>%
  mutate_at(vars(-drug1), function(x) round(x, 2)) %>%
  gather('drug2', 'corr', -drug1) %>%
  filter(drug1 != drug2) %>%
  mutate_at(vars(drug1, drug2), drug_factor) %>%
  ggplot(aes(y=drug1, x=drug2, fill=corr)) +
  geom_tile() +
  scale_fill_distiller(palette='RdGy') +
  theme_classic()
```

# Discussion

Notably, all the correlations are positive: taking one drug means you are more likely to take another.

The strongest correlations are between ciprofloxacin and each of metronidazole, nitrofurantoin, and TMP/SMX. TMP/SMX and nitrofurantoin also correlate.

Levofloxacin has a weaker correlation with metronidazole than ciprofloxacin does, but azithromycin's strongest correlation is with azithromycin.