### 2017-03-28 top-users/top-users.Rmd
more tables

---
title: "Top antibiotic users"
author: "Scott Olesen"
date: "27 March 2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, 
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

```{r}
pn = function(x) prettyNum(x, big.mark=',')
```

# Motivation

What drugs are the top users taking? Are they different from the drugs that people take in general? Are the patterns of drug-drug correlations different among the top users?

# Data

I use the 2011 Medicare data. I will sometimes exclude claims for the anti-tuberculosis drugs (ethambutol, rifampin, isoniazid, pyrazinamide) because many of the "top" users, depending on the defintion, are taking these drugs, which constitute a different problem.

I merged amoxicillin/clavulanate into amoxicillin.

```{r load_data}
bene = read_tsv('../bene_2011.tsv')
n_bene = nrow(bene)
pde = read_tsv('../pde_2011.tsv') %>%
  rename(drug=antibiotic) %>%
  mutate(drug=if_else(drug=='amoxicillin/clavulanate', 'amoxicillin', drug)) %>%
  group_by(bene_id, drug) %>%
  summarize(n_claims=n(), n_days=sum(days_supply)) %>%
  ungroup()
```

# Results

## Drug consumption sort-of follows a power law

That is, the *exponent* of the number of beneficiaries with a certain of claims *n* (or days' supply of drug) falls with *n*.

```{r drug_distribution}
pde_total = pde %>%
  group_by(bene_id) %>%
  summarize(n_total_claims=sum(n_claims), n_total_days=sum(n_days))

pde_total %>%
  count(n_total_claims) %>%
  ggplot(aes(x=n_total_claims, y=n)) +
  geom_point() +
  scale_y_log10() +
  xlab('number of claims') + ylab('number of beneficiaries') +
  theme_minimal()

pde_total %>%
  count(n_total_days) %>%
  ggplot(aes(x=n_total_days, y=n)) +
  geom_point() +
  scale_y_log10() +
  xlab('number of days\' supply') + ylab('number of beneficiaries') +
  theme_minimal()
```

## Top users by claim

- Merge amox, amox/clav
- Exclude TB drugs? Do with and without?
- Split people into top users (by # claims) or not
- Analyze them that way (with CPKP)
- Add zeros for people/drug combos? Or just find one denominator and pad out
- Then repeat the same thing but using days (and DID)
- Make plots & tables to compare both

Expected punchlines:

- TB drugs make up a lot of the top use
- Top claims enrich for vancomycin, which people are taking with ~2 day claims
- Top days are for mostly the same drugs

```{r top_claims_use}
top_claims_cutoff = 15

# classify beneficiaries as in top bracket or not
top_claims_bene = pde %>%
  group_by(bene_id) %>%
  summarize(top = sum(n_claims) >= top_claims_cutoff)

n_top_claims_bene = sum(top_claims_bene$top)

top_claims_table = pde %>%
  left_join(top_claims_bene, by='bene_id') %>%
  group_by(drug, top) %>%
  summarize(n_claims=sum(n_claims)) %>%
  ungroup() %>%
  mutate(n_bene=if_else(top, n_top_claims_bene, n_bene - n_top_claims_bene),
         cpkp=n_claims*1000/n_bene) %>%
  select(drug, top, cpkp) %>%
  spread(top, cpkp) %>%
  arrange(desc(`TRUE`)) %>%
  head(20) %>%
  rename(`non-top user CPKP`=`FALSE`, `top user CPKP`=`TRUE`) %>%
  kable
```

Of `r pn(n_bene)` total beneficiaries, there are `r pn(n_top_claims_bene)` that had more than `r top_claims_cutoff` claims. In aggregate, their consumption looks fairly different from the rest of the population:

`r top_claims_table`

Notable differences include:

- More urinary anti-infectives (Bactrim, nitro, and trimethoprim)
- Much more vancomycin
- More anti-TB drugs (esp. rifampin, ethambutol)

## Top users by days

Measuring top users by claims biases toward people taking drugs that are given with short dosages.

```{r top_days_use}
# classify beneficiaries as in top bracket or not
top_days_bene = pde %>%
  group_by(bene_id) %>%
  summarize(top = sum(n_days) >= 365)

n_top_days_bene = sum(top_days_bene$top)

top_days_table = pde %>%
  left_join(top_days_bene, by='bene_id') %>%
  group_by(drug, top) %>%
  summarize(n_days=sum(n_days)) %>%
  ungroup() %>%
  mutate(n_bene=if_else(top, n_top_days_bene, n_bene - n_top_days_bene),
         did=n_days*1000/(365*n_bene)) %>%
  select(drug, top, did) %>%
  spread(top, did) %>%
  arrange(desc(`TRUE`)) %>%
  head(20) %>%
  rename(`non-top user DID`=`FALSE`, `top user DID`=`TRUE`) %>%
  kable
```

Of `r pn(n_bene)` total beneficiaries, there are `r pn(n_top_days_bene)` that had more than 365 days of days' supply. In aggregate, their consumption looks fairly different from the rest of the population:

`r top_days_table`

The drugs that make it into the top-days-users top list are similar to those from the top-claims-users, except that vancomycin is missing here.

## Top claims, no anti-TB

I worry that we're mostly selecting for people who have TB when doing these analyses, so I did a different version where I excluded the TB drugs:

```{r tb_drugs}
tb_drugs = c('isoniazid', 'rifabutin', 'rifampin', 'rifampin/isoniazid', 'rifampin/isoniazid/pyrazinamide', 'rifapentine', 'rifamixin', 'pyrazinamide', 'ethambutol')
stopifnot(all(tb_drugs %in% unique(pde$drug)))

data_frame(drug=tb_drugs) %>% kable
```

```{r top_claims_notb}
top_notb_cutoff = 15

# classify beneficiaries as in top bracket or not
top_notb_bene = pde %>%
  filter(!(drug %in% tb_drugs)) %>%
  group_by(bene_id) %>%
  summarize(top = sum(n_claims) >= top_notb_cutoff)

n_top_notb_bene = sum(top_notb_bene$top)

top_notb_table = pde %>%
  left_join(top_notb_bene, by='bene_id') %>%
  group_by(drug, top) %>%
  summarize(n_claims=sum(n_claims)) %>%
  ungroup() %>%
  mutate(n_bene=if_else(top, n_top_notb_bene, n_bene - n_top_notb_bene),
         cpkp=n_claims*1000/n_bene) %>%
  select(drug, top, cpkp) %>%
  spread(top, cpkp) %>%
  arrange(desc(`TRUE`)) %>%
  head(20) %>%
  rename(`non-top user CPKP`=`FALSE`, `top user CPKP`=`TRUE`) %>%
  kable
```

```{r compute_user_overlap}
# The idea here was to see about the overlap between drug users.
# If you took drug 1, how likely was it for you to also take drug 2?
# This is actually pretty similar to the tetrachoric analysis I did previously.
# Instead, I summarized the confusion matrix with the Matthews correlations coefficient.

drugs = pde$drug %>% unique

user_overlap = function(pde, drug1_name, drug2_name) {
  full_join(pde %>% filter(drug==drug1_name) %>% rename(drug1=drug),
            pde %>% filter(drug==drug2_name) %>% rename(drug2=drug), by='bene_id') %>%
    replace_na(list(drug1=0, drug2=0)) %>%
    mutate(case=case_when(.$drug1 > 0 & .$drug2 > 0 ~ 'took_both',
                          .$drug1 > 0 ~ 'took_drug1_only',
                          .$drug2 > 0 ~ 'took_drug2_only',
                          TRUE ~ 'took_neither')) %>%
    count(case) %>%
    spread(case, n) %>%
    mutate(drug1=drug1_name, drug2=drug2_name)
}

mcc = function (tp, fp, fn, tn) {
  # tp = true positive, fp = false positive, etc.
  n = as.double(tp + fp + fn + tn)
  s = (tp + fn) / n
  p = (tp + fp) / n
  return (tp / n - (s * p)) / sqrt(p * s * (1 - s) * (1 - p))
  #return ((tp * tn) - (fp * fn)) / sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
}

#drug_overlap = crossing(drug1=drugs, drug2=drugs) %>%
#  filter(drug1 > drug2) %>%
#  rowwise() %>% do(user_overlap(pde, .$drug1, .$drug2)) %>%
#  mutate(took_neither=n_bene-(took_both + took_drug1_only + took_drug2_only)) %>%
#  select(drug1, drug2, took_both, took_drug1_only, took_drug2_only, took_neither) %>%
#  mutate(mcc=mcc(took_both, took_drug1_only, took_drug2_only, took_neither))
```

```{r top_user_overlap}
# The idea here was to ask which users are the "top" users for that drug.
# Then I would overlap them in the same way that I did for any-users above.

pad_to = function(x, n, with=0) {
  if (length(x) > n) stop('input longer than pad to length')
  c(x, rep(0, n - length(x)))
}

top_bracket_limit = function(n, fraction) {
  n[cume_dist(n) > (1 - fraction)] %>% min
}

#drug_upper_brackets = pde %>%
#  inner_join(pde %>% group_by(bene_id) %>% summarize(n_claims=sum(n_claims)) %>% mutate(antibiotic='all'), by='bene_id') %>%
#  group_by(antibiotic) %>%
#  summarize(limit=top_bracket_limit(pad_to(n_claims, n_bene), 0.001)) %>%
#  filter(limit > 0)
```