### 2017-04-04 dosage-form/dosage-form.Rmd
streamlined dosage

---
title: "Dosage forms in antibiotic PDEs"
author: "Scott Olesen"
date: "2/20/2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4,
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

```{r load_formulary, message=FALSE}
formulary = 2011:2014 %>%
  lapply(function(y) read_tsv(sprintf('../../data/formulary/formulary_%i.tsv', y)) %>%
           select(brand_name=BN, generic_name=GNN, dosage_form_code=GCDF, dosage_form_desc=GCDF_DESC) %>%
           mutate(year=y)) %>%
  bind_rows()
```

I no longer have the raw PDE data, so I can't use that!
```{r load_pde, message=FALSE}
# Start with just one year
#pde = 2011:2012 %>%
  #lapply(function(y) read_tsv(sprintf('../pde_%i.tsv', y)) %>% mutate(year=y)) %>%
  #bind_rows()
```

Most codes have a unique descriptions.
```{r code_desc}
unique_combos = formulary %>%
  select(dosage_form_code, dosage_form_desc) %>%
  distinct %>%
  arrange(dosage_form_code)

unique_combos %>% kable
```

There are a few exceptions:

```{r repeated_codes}
unique_combos %>%
  group_by(dosage_form_code) %>%
  filter(length(dosage_form_desc) > 1) %>%
  arrange(dosage_form_code, dosage_form_desc) %>%
  kable
```

None of these repetitions are meaningfully different from one another, so I'll
conclude that the code uniquely identifies the dosage type.

```{r codes_by_year}
#canonical_codes = unique_combos %>%
#  group_by(dosage_form_code) %>%
#  do(head(., 1))
#
#dat %>%
#  select(dosage_form_code, year) %>%
#  group_by(year) %>%
#  count(dosage_form_code) %>%
#  spread(year, n) %>%
#  left_join(canonical_codes, by='dosage_form_code') %>%
#  arrange(desc(`2011`)) %>%
#  kable
```

Which ones of the codes should I keep? Let's consider, for each code, the top drugs that match that code. Of these, I should only keep some.

```{r drugs_by_code, message=FALSE}
#drugs_by_code = dat %>%
#  filter(year == 2011) %>%
#  select(generic_name, brand_name, dosage_form_code) %>%
#  group_by(dosage_form_code) %>%
#  count(generic_name) %>%
#  group_by(dosage_form_code) %>%
#  do(head(., 3))

codes_to_keep = scan('oral_injected_codes.txt', what=character())

#drugs_by_code %>%
#  mutate(keep_code=dosage_form_code %in% codes_to_keep) %>%
#  kable
```

For posterity, I export a list of the code/values that I kept.
```{r save_codes}
unique_combos %>%
  filter(dosage_form_code %in% codes_to_keep) %>%
  arrange(dosage_form_code) %>%
  write_tsv('code_forms.tsv')
```