### 2017-05-10 temporal/seasonal-traces/drug_seasonality.Rmd
improved seasonality

---
title: "Antibiotic seasonalities"
author: "Scott Olesen"
date: "5/10/2017"
output: html_document
---

# Background

For the GISP project, we are trying to show that we can infer bystander selection pressure on gonococcus using the seasonality of consumption of drugs that are consumed for reasons other than gonorrhea treatment.

# Methods

## Data

I use the 2011--2014 Medicare data, filtered for:

- beneficiaries between 66 and 96
- beneficiaries without any HMO months
- beneficiaries with a non-NA state
- with a sex in male/female

## Analysis

I use the `spec.pgram` function with `detrend=TRUE`, which removes a linear trend from the whole data (and also removes the mean).

I also use `nls` to fit the curves. The rough equation is $y = A + Bt + C\sin(2\pi t/D + E\pi)$. The time $t$ is just weeks, and $y$ is the number of claims in that week. This function works only if you give good starting values:

- vertical offset $A$ as the mean number of claims
- secular trend slope $B \approx 0$
- amplitude $C$ as the standard deviation of the number of claims
- period $D \approx 52$ (because we expect a yearly period)
- phase offset $E \approx \tfrac{1}{2}$ (because we expect consumption to be high in the winter when $t \approx 0$)

For the plots, I show a fit that includes a linear

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
                      pdf.options(useDingbats=FALSE),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
import::from(broom, tidy)
```

```{r load_data}
dat = lapply(2011:2014, function(y) {
  sprintf('../temporal_abx_%i.tsv', y) %>%
    read_tsv %>%
    filter(!is.na(state), week <= 52) %>%
    mutate(year=y)
  }) %>%
  bind_rows %>%
  rename(week_of_year=week) %>%
  mutate(week=(year - 2011)*52 + week_of_year)
```

```{r spectrogram_functions, echo=TRUE}
strongest_period = function(pg) 1.0 / (pg$freq[which.max(pg$spec)])
drug_data = function(drugs) {
  filter(dat, antibiotic %in% drugs) %>%
    group_by(week) %>%
    summarize(n_claims=sum(n_claims))
}

drug_plot = function(drugs) {
  pf = periodic_fit(drugs)
  p = pf$m$getPars()
  f = function(t) p['A'] + p['B'] * t + p['C'] * sin(2 * pi * t / p['D'] + pi * p['E'])
  drug_data(drugs) %>%
    ggplot(aes(x=week, y=n_claims)) + geom_point() + stat_function(fun=f)
}

drug_spectrogram = function(drugs) drug_data(drugs) %$% spec.pgram(n_claims)
drug_period = function(drugs) strongest_period(drug_spectrogram(drugs))

periodic_fit = function(drugs) {
  d = drug_data(drugs) %>% mutate(x=week, y=n_claims)
  nls(y ~ A + B*x + C*sin(2*pi*x/D + pi*E), start=list(A=mean(d$y), B=0, C=sd(d$y), D=52, E=0.5), data=d)
}

periodic_fit_table = function(drugs) periodic_fit(drugs) %>% tidy %>% select(term, estimate) %>% kable
```

# Results

Azithromycin and quinolones are strongly seasonal: the spectrogram has a clear peak at approximately 52 week period, and the curve fit gives a strong amplitude and a period of almost exaclty 52. The quinolones are also seasonal---there is a peak in the spectrogram and the curve fit works nicely---but the amplitude is not as strong.

The tetracyclines are moderately seasonal: the spectrogram and fit both work, but the amplitudes are again smaller. Ceftriaxone is weakly seasonal: the spectrogram does find that approximately 52 weeks is the strongest period, but the curve fit surprises me. The penicillins are not seasonal at all.

## Azithromycin

```{r azithromycin}
drugs = c('azithromycin')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
periodic_fit_table(drugs)
```

The strongest period in the spectrogram is `r drug_period(drugs)` weeks.

## Cipro/levo

```{r quinolone}
drugs = c('levofloxacin', 'ciprofloxacin')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
periodic_fit_table(drugs)
```

The strongest period in the spectrogram is `r drug_period(drugs)` weeks.

## Tetracycline/doxycycline

```{r tetra}
drugs = c('tetracycline', 'doxycycline')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
periodic_fit_table(drugs)
```

The strongest period in the spectrogram is `r drug_period(drugs)` weeks.

## Ceftriaxone

```{r ceftriaxone}
drugs = c('ceftriaxone')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
periodic_fit_table(drugs)
```

The strongest period in the spectrogram is `r drug_period(drugs)` weeks.

## Pencillin/ampicillin

```{r penicillin}
drugs = c('penicillin', 'ampicillin')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
periodic_fit_table(drugs)
```

The strongest period in the spectrogram is `r drug_period(drugs)` weeks.