### 2017-06-21 ms2/ms2.Rmd
moving to HRR

---
title: "ms2"
author: "Scott Olesen"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
                      pdf.options(useDingbats=FALSE),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
import::from(broom, tidy)
import::from(tools, md5sum)
```

```{r load_data}
regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
  select(state, state_abbreviation, region)

zipcodes = read_tsv('../../db/zipcode/zipcode.tsv') %>%
  select(zipcode=zip, state)

# Use the zipcode-HRR mapping from 2011
hrr = read_tsv('../../db/hrr/hrr.tsv') %>%
  filter(year==2011) %>%
  select(zipcode, hrr) %>%
  left_join(zipcodes, by='zipcode') %>%
  left_join(regions, by='state')

years = 2011:2014
read_years = function(template_fn) {
  lapply(years, function(y) read_tsv(sprintf(template_fn, y)) %>% mutate(year=y)) %>%
    bind_rows()
}

all_ineq = read_years('data/state_ineq_all_%i.tsv') %>%
  group_by(state) %>%
  summarize(mean_cpb=mean(mean),
            sem_cpb=sd(mean)/sqrt(n()),
            mean_fnz=mean(fnz),
            sem_fnz=sd(fnz)/sqrt(n()),
            mean_nzgini=mean(nzgini),
            sem_nzgini=sd(nzgini)/sqrt(n())) %>%
  left_join(regions, by='state')

state_ineq = read_years('data/state_ineq_%i.tsv') %>%
  rename(cpb=mean, nzg=nzgini, nzr=mean_fnz_residual) %>%
  group_by(state, drug_group) %>%
  summarize_at(vars(-state, -drug_group, -total), mean) %>%
  ungroup()

hrr_ineq = read_years('data/hrr_ineq_%i.tsv') %>%
  rename(cpb=mean, nzg=nzgini, nzr=mean_fnz_residual) %>%
  group_by(hrr, drug_group) %>%
  summarize_at(vars(-hrr, -drug_group, -total), mean) %>%
  ungroup()
```

```{r load_abg, cache.extra=md5sum('bug_drug.tsv')}
resistance_groups = read_tsv('resistance_groups.tsv')
bug_names = read_tsv('bug_names.tsv')
bug_groups = read_tsv('bug_groups.tsv')
possible_bug_drug = read_tsv('bug_drug.tsv')

raw_abg = read_tsv('../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
  mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
  left_join(hrr, by=c('zipcode', 'state')) %>%
  right_join(resistance_groups, by='drug') %>%
  select(bug, raw_drug=drug, drug_group, state, hrr, percent_nonsusceptible, n_isolates)

# Keep only the abundant bug/drug combinations
# min_n_states = 35
# good_combos = semi_join(raw_abg, possible_bug_drug, by=c('bug', 'drug_group')) %>%
#   select(bug, drug_group, state) %>%
#   distinct %>%
#   count(bug, drug_group) %>% ungroup() %>% 
#   rename(n_places=n) %>%
#   filter(n_places >= min_n_states)

abg = raw_abg %>%
  #mutate(raw_bug=bug) %>%
  # shorten the bug names
  left_join(bug_names, by='bug') %>%
  select(-bug) %>% rename(bug=bug_short) %>%
  # group the different S. aureus into meta-groups depending on the drug
  left_join(bug_groups, by=c('bug', 'drug_group')) %>%
  mutate(bug_group=if_else(is.na(bug_group), bug, bug_group)) %>%
  select(-bug) %>% rename(bug=bug_group) %>%
  #right_join(good_combos, by=c('bug', 'drug_group')) %>%
  # keep only particular bug/drug combos
  right_join(possible_bug_drug, by=c('bug', 'drug_group'))
  #filter(!is.na(percent_nonsusceptible), !is.na(n_isolates))

# get the median number of isolates for each bug/drug
median_n_isolates = abg %>%
  filter(!is.na(n_isolates)) %>%
  group_by(bug, drug_group) %>%
  summarize(median_n_isolates=as.integer(median(n_isolates))) %>%
  ungroup()

# get the number of states in which each bug/drug appears
min_n_states = 37
abg_n_states = abg %>%
  select(bug, drug_group, state) %>%
  distinct() %>%
  count(bug, drug_group) %>%
  ungroup() %>%
  rename(n_states=n)
abg_n_hrr = abg %>%
  select(bug, drug_group, hrr) %>%
  distinct() %>%
  count(bug, drug_group) %>%
  ungroup() %>%
  rename(n_hrr=n)

# replace the NA number of isolates with the median for that bug/drug
abg %<>%
  left_join(median_n_isolates) %>%
  mutate(n_isolates=if_else(is.na(n_isolates), median_n_isolates, n_isolates)) %>%
  select(-median_n_isolates) %>%
  left_join(abg_n_states, by=c('bug', 'drug_group')) %>%
  left_join(abg_n_hrr, by=c('bug', 'drug_group'))
  #filter(n_places >= min_n_states)

# check that all the short names made it in
for (x in c('bug', 'n_isolates', 'percent_nonsusceptible')) {
  if (any(is.na(abg[[x]]))) stop(sprintf('NAs in field "%s"', x))
}

summarize_abg = function(df, place_name) {
    group_by_(df, 'bug', 'drug_group', place_name) %>%
    summarize(mean_percent_nonsusceptible=weighted.mean(percent_nonsusceptible, sqrt(n_isolates)),
              n_place_antibiograms=n()) %>%
    ungroup() %>%
    mutate(place_weight=sqrt(n_place_antibiograms))
}

state_abg = summarize_abg(abg, 'state')
hrr_abg = summarize_abg(abg, 'hrr')
```

```{r do_models}
linear_interval = function(model, level) {
  confint(model, level=level) %>%
    tidy %>%
    setNames(c('term', 'ci_low', 'ci_high'))
}

linear_model = function(df, dependent_var, explanatory_vars) {
  f = paste0(dependent_var, ' ~ ', paste0(explanatory_vars, collapse = ' + ')) %>% formula
  m = lm(formula=f, weights=place_weight, data=df)
  tidy(m) %>%
    select(term, estimate, p.value) %>%
    mutate(n=nrow(df)) %>%
    left_join(linear_interval(m, 0.95), by='term')
}

spearman_interval = function(df, rho_name, n_name, level) {
  mada::CIrho(df[[rho_name]], df[[n_name]], level=level) %>%
    tidy %>%
    setNames(c('rho', 'ci_low', 'ci_high')) %>%
    select(-rho)
}

spearman_model = function(df, dependent_var, explanatory_var) {
  cor.test(df[[dependent_var]], df[[explanatory_var]], method='spearman') %>%
    tidy %>%
    mutate(term=explanatory_var, method='spearman') %>%
    mutate(n=nrow(df)) %>%
    group_by(term, estimate, p.value, n) %>%
    do(spearman_interval(., 'estimate', 'n', 0.95)) %>%
    ungroup()
}

uni_term = 'cpb' # univariable explanatory term
multi_term_adjuster = 'cpb'
multi_term = 'nb_prob' # multivariable explanatory term (along with fnz)
models = function(x) {
  dep = 'mean_percent_nonsusceptible'
  bind_rows(
    linear_model(x, dep, uni_term) %>% mutate(model='uni'),
    linear_model(x, dep, c(multi_term_adjuster, multi_term)) %>% mutate(model='multi'),
    spearman_model(x, dep, uni_term) %>% mutate(model='spearman')
  ) %>%
    mutate(n_data=nrow(x))
}

results = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  group_by(bug, drug_group) %>%
  do(models(.)) %>%
  ungroup

hrr_results = left_join(hrr_abg, hrr_ineq, by=c('hrr', 'drug_group')) %>% 
  group_by(bug, drug_group) %>%
  do(models(.)) %>%
  ungroup
```

# Landscape of bug/drug correlations

## Selecting correlations to test

I join the resistance drugs into classes:

`r kable(resistance_groups)`

I similarly joined the consumption drugs into similar classes.

I started by considering these `r nrow(possible_bug_drug)` bug/drug combinations on the basis of their clinical relevance and plausibility:

`r kable(possible_bug_drug)`

In previous iterations, I filtered by the number of states in which each bug/drug combination appeared. Increasing that threshold ends up testing the relationships for which we have the best data, and therefore we are the best powered for. Unfortunately, my simulations mostly make it seem like we should, to get the best power, use all the bug/drug combinations.

```{r, eval=FALSE}
#Then I filtered to keep only those bug/drug combinations that were present in `r min_n_states` states. That means I'm looking at these `r nrow(good_combos)` bug/drug combinations:

#`r good_combos %>% arrange(desc(n_places)) %>% kable`
```

## Preparing the consumption data

Unlike previous analyses, this time I treated the consumption and resistance data similarly: I averaged both across all data years. Thus, to get a state's Gini coefficient, I computed that number each year and then took the average.

## Spearman

Significant results using Spearman correlations of mean percent nonsusceptible ~ claims per beneficiary. A star after the bug/drug combination implies that the adjusted $p < 0.05$, and a dot means the unadjusted $p < 0.05$.

```{r spearman_models}
results %>%
  filter(model=='spearman') %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  filter(p.value < 0.05) %>%
  arrange(p.adj) %>%
  select(bug, drug_group, estimate, p.value, p.adj, n) %>%
  kable
```

Aside from individual significant results, we can look at the landscape of correlations. They are almost all positive, with MRSA/Bactrim making an interesting exception. Lines show 95% confidence intervals.

```{r spearman_interval}
interval_plot = function(df) {
  arrange(df, estimate) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  mutate(sig=case_when(.$p.adj < 0.05 ~ '*',
                       .$p.value < 0.05 ~ '.',
                       TRUE ~ ''),
         bug_drug=interaction(bug, drug_group, sig),
         bug_drug=factor(bug_drug, levels=bug_drug, ordered=TRUE)) %>% 
    ggplot(aes(x=bug_drug, y=estimate, ymin=ci_low, ymax=ci_high, color=drug_group)) +
    geom_point() +
    geom_linerange() +
    coord_flip()
}

results %>%
  filter(model=='spearman') %>%
  interval_plot() +
  ylab("Spearman's rho")
```

HRR spearman:

```{r hrr_spearman_interval}
hrr_results %>%
  filter(model=='spearman') %>%
  interval_plot() +
  ylab("Spearman's rho")
```

```{r, eval=FALSE}
# Used this to show that there's a correlation between
# strength of correlation and amount of data
x = abg %>% 
  select(bug, drug_group, state) %>%
  distinct() %>%
  count(bug, drug_group) %>% ungroup() %>% rename(n_places=n)
results %>% filter(model=='spearman') %>%
  left_join(x, by=c('bug', 'drug_group')) %>%
  ggplot(aes(x=n_places, y=estimate)) +
  geom_point()
```

## Differences in consumption/resistance patterns explain some of these results

Different bug/drug combinations have different spreads along the consumption and resistance axes, so we would expect to see different strengths of association for each one.

```{r}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  filter(bug %in% c('E. coli', 'Eb. cloacae', 'CoNS', 'MSSA', 'MRSA', 'S. pneumoniae')) %>%
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, color=drug_group, shape=bug)) +
  geom_point() +
  xlab('claims per beneficiary')
```

# Unevenness of consumption and antibiotic resistance

## Drugs are consumed differently in different places

```{r nzgini_cpb_drug}
cbbPalette <- c("#000000", "#D55E00", "#CC79A7", "#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")
state_ineq %>%
  ggplot(aes(x=cpb, y=nzg, color=drug_group)) +
  geom_point() +
  ylab('Gini coefficient among users') + xlab('claims per beneficiary') +
  theme_minimal() +
  scale_color_manual(values=cbbPalette)
```

## Effect of unevenness on antibiotic resistance

Here we use the same bug/drug combinations, but instead we test the multivariable model mean percent nonsusceptible ~ fraction nonzero + nonzero Gini.

There is only one significant and one marginally significant result:

```{r multi_models}
results %>%
  filter(model=='multi', term==multi_term) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  filter(p.value < 0.05) %>%
  arrange(p.value) %>%
  select(bug, drug_group, estimate, p.value, p.adj) %>%
  kable
```

```{r}
results %>%
  filter(model=='multi', term==multi_term) %>%
  interval_plot() +
  ylab('Gini coefficient')
```

Using HRRs, one of the answers remains the same, but a different marginally-significant result pops up.

```{r}
hrr_results %>%
  filter(model=='multi', term==multi_term) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  filter(p.value < 0.05) %>%
  arrange(p.value) %>%
  select(bug, drug_group, estimate, p.value, p.adj) %>%
  kable
```

```{r}
hrr_results %>%
  filter(model=='multi', term==multi_term) %>%
  interval_plot()
```

## Specific plots

Quinolone consumption predicts *E. coli* quinolone resistance:

```{r ecoli_quin}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='E. coli', drug_group=='quinolone') %>%
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  xlim(0.2, 0.45) + ylim(0, 50) +
  xlab('quinolone claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
  theme_minimal()
```

## E.coli and bactrim

```{r ecoli_bactrim}
ecoli_bactrim_dat = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') #%>%
  #filter(!is.na(mean_percent_nonsusceptible))

ecoli_bactrim_dat %>%
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
  ylim(10, 45) +
  xlab('TMP/SMX claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
  theme_minimal()
```

```{r}
ecoli_bactrim_dat %>%
  ggplot(aes(x=nzg, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  xlab('Gini coefficient)') +
  ylab('E. coli resistance') +
  theme_minimal()
```

## Pneumococcus and macrolides

```{r}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='S. pneumoniae', drug_group=='macrolide') %>%
  filter(!is.na(mean_percent_nonsusceptible)) %>% 
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  theme_minimal()
```

## E. cloacae and quinolones

```{r}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='Eb. cloacae', drug_group=='quinolone') %>%
  filter(!is.na(mean_percent_nonsusceptible)) %>%
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  theme_minimal()
```

## S. aureus and clindamycin

```{r staph_clinda}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='S. aureus', drug_group=='clindamycin') %>%
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  xlab('clindamycin claims per beneficiary') + ylab('mean percent nonsusceptible S. aureus isolates') +
  theme_minimal()
```

# Supplementary results

## Metrics of unevenness

### Aggregate usage correlates with fraction of users

Error bars show the s.e.m. across the four data years.

```{r fnz_mean}
all_ineq %>%
  mutate_at(vars(matches('fnz')), function(x) x*100) %>%
  ggplot(aes(x=mean_cpb, xmin=mean_cpb-sem_cpb, xmax=mean_cpb+sem_cpb,
             y=mean_fnz, ymin=mean_fnz-sem_fnz, ymax=mean_fnz+sem_fnz,
             color=region, label=state_abbreviation)) +
  geom_point() +
  geom_linerange() +
  geom_errorbarh(height=0) +
  geom_text(nudge_x=0.04) +
  xlab('mean claims per beneficiary') +
  ylab('mean % beneficiaries with at least one claim') +
  theme_minimal()
```

```{r}
all_ineq %$%
  cor.test(mean_cpb, mean_fnz, method='spearman')
```

### The two inequality metrics not correlate well

Error bars show the s.e.m. across the four data years.

```{r nzgini_mean}
all_ineq %>%
  ggplot(aes(x=mean_fnz, xmin=mean_fnz-sem_fnz, xmax=mean_fnz+sem_fnz,
             y=mean_nzgini, ymin=mean_nzgini-sem_nzgini, ymax=mean_nzgini+sem_nzgini,
             color=region, label=state_abbreviation)) +
  geom_point() +
  geom_linerange() +
  geom_errorbarh(height=0) +
  geom_text(nudge_x=0.00) +
  xlab('mean fraction with at least one claim') +
  ylab('mean Gini coefficient') +
  theme_minimal()
```

The correlation exists but it's much less strong.

```{r}
all_ineq %$%
  cor.test(mean_fnz, mean_nzgini, method='spearman')
```

There is an intermediate correlation between claims per beneficiary and the Gini coefficient (see below), so in multivariable regressions I will use those two (rather than CPB and Gini).

```{r}
all_ineq %$% cor.test(mean_cpb, mean_nzgini, method='spearman')
```

## Spearman models

### Comparing state and HRR

HRR-level Spearman correlations between consumption and resistance (almost) always underestimate the correlation reported at the state level:

```{r}
bind_rows(mutate(results, group='state'),
          mutate(hrr_results, group='hrr')) %>%
  filter(model=='spearman') %>%
  arrange(estimate) %>%
  mutate(bug_drug=interaction(bug, drug_group)) %>%
  select(bug_drug, group, estimate) %>%
  spread(group, estimate) %>%
  ggplot(aes(x=state, y=hrr, label=bug_drug)) +
  geom_point() +
  geom_text() +
  geom_abline()
```

### State-level volcano plot

```{r}
results %>%
  filter(model=='spearman') %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  ggplot(aes(x=estimate, y=-log10(p.adj))) + geom_point() +
  geom_hline(yintercept=-log10(0.05), col='red')
```

### HRR-level correlations

```{r hrr_spearman_models}
hrr_results %>%
  filter(model=='spearman') %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  filter(p.value < 0.05) %>%
  arrange(p.value) %>%
  select(bug, drug_group, estimate, p.value, p.adj) %>%
  kable
```

## Univariable regressions

Like the Spearman models, the model here is mean percent nonsusceptible ~ claims per beneficiary, only this time it's a normal linear regression.

### State-level

```{r uni_models}
results %>%
  filter(model=='uni', term==uni_term) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  #filter(p.value < 0.05) %>%
  arrange(p.value) %>%
  select(bug, drug_group, estimate, p.value, p.adj, n) %>%
  kable
```

```{r}
results %>%
  filter(model=='uni', term==uni_term) %>%
  interval_plot()
```

```{r}
results %>%
  filter(model=='uni') %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  ggplot(aes(x=estimate, y=-log10(p.adj))) + geom_point() +
  geom_hline(yintercept=-log10(0.05), col='red')
```

### HRR-level

```{r hrr_uni_models}
hrr_results %>%
  filter(model=='uni', term==uni_term) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  filter(p.value < 0.05) %>%
  arrange(p.value) %>%
  select(bug, drug_group, estimate, p.value, p.adj) %>%
  kable
```