### 2017-06-22 sex-difference/namcs.Rmd
namcs sex diff

---
title: "Drivers of sex differences in consumption (NAMCS)"
author: "Scott Olesen"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)

import::from(survey, svyglm)
import::from(survey, svydesign)
```

```{r load_data}
sex_codes = read_tsv('namcs_sex_codes.tsv') %>%
  rename(dx=code)

dx = read_tsv('../../db/namcs/data/namcs_2013_dx.tsv') %>%
  left_join(sex_codes, by='dx')

dx_visit = dx %>%
  group_by(record_id) %>%
  summarize(uti='uti' %in% diagnosis_type,
            skin='skin' %in% diagnosis_type,
            vaginitis='vaginitis' %in% diagnosis_type)

abx_drug_codes = read_tsv('../../db/namcs/drug-codes/abx-codes.tsv') %>%
  select(drug_code, antibiotic) %>%
  mutate(antibiotic=tolower(antibiotic))

drugs = read_tsv('../../db/namcs/data/namcs_2013_drugs.tsv')

abx = drugs %>% left_join(abx_drug_codes, by='drug_code') %>% filter(!is.na(antibiotic))

abx_visits = abx %>%
  select(antibiotic, record_id) %>%
  distinct() %>%
  mutate(value=TRUE) %>% spread(antibiotic, value)

visits = read_tsv('../../db/namcs/data/namcs_2013_visits.tsv') %>%
  left_join(dx_visit, by='record_id') %>%
  replace_na(list(uti=FALSE, skin=FALSE, vaginitis=FALSE)) %>%
  left_join(abx_visits, by='record_id') %>%
  mutate_at(vars(amoxclav:tmp_smx), function(x) !is.na(x))

abx_names = visits %>% select(amoxclav:tmp_smx) %>% names
stopifnot(length(abx_names) == length(unique(abx_drug_codes$antibiotic)))
```

Most antibiotics are prescribed in the context of a new (less than 3 month onset) problem, but the people who get antibiotics in the context of a chronic problem tend to be older:

```{r}
abx %>%
  select(antibiotic, record_id) %>% distinct %>%
  left_join(visits, by='record_id') %>%
  group_by(major_reason) %>%
  summarize(n=n(), mean_age=mean(age), sd_age=sd(age)) %>%
  arrange(desc(n)) %>%
  kable
```

# Overall

```{r}
visits %>%
  right_join(abx, by='record_id') %>%
  group_by(sex, antibiotic) %>%
  summarize(n_rx=sum(weight)) %>%
  ungroup() %>%
  spread(sex, n_rx) %>%
  ggplot(aes(x=male, y=female, label=antibiotic)) +
  geom_point() +
  geom_text() +
  coord_fixed() +
  geom_abline()
```

```{r models}
des = visits %>%
  mutate(sex=factor(sex, levels=c('male', 'female'))) %>%
  svydesign(data=., id=~0, weights=~weight)

f = function(drug) {
  svyglm(as.formula(sprintf('%s ~ sex', drug)), des, family=binomial(link='log'), start=c(-5, 0)) %>%
    tidy %>%
    filter(term=='sexfemale') %>%
    mutate(drug=drug)
}

sex_diff_by_drug = lapply(abx_names, f) %>%
  bind_rows() %>%
  mutate(ci=1.96*std.error,
         ci_low=estimate-ci,
         ci_hi=estimate+ci,
         p.adj=p.adjust(p.value, method='BH')) %>%
  mutate(sig=case_when(.$p.adj < 0.05 ~ '**',
                       .$p.value < 0.05 ~ '*',
                       TRUE ~ '')) %>%
  select(-ci) %>%
  mutate_at(vars(estimate, ci_low, ci_hi), function(x) exp(x))
```

```{r diff_plot}
sex_diff_by_drug %>%
  mutate(drug=forcats::fct_reorder(factor(drug), estimate)) %>%
  ggplot(aes(x=drug, y=estimate, ymin=ci_low, ymax=ci_hi)) +
  geom_point() + geom_linerange() +
  geom_text(aes(y=1, label=sig), cex=6) +
  coord_flip() +
  scale_y_continuous(breaks=c(0, 1, 2, 5, 10), limits=c(0, 10)) +
  ylab('relative risk of being female for getting drug')
```

```{r diff_table}
sex_diff_by_drug %>%
  select(drug, ci_low, estimate, ci_hi, p.value, p.adj, sig) %>%
  mutate_at(vars(ci_low:ci_hi), function(x) round(x, digits=2)) %>%
  mutate_at(vars(p.value, p.adj), function(x) round(x, digits=3)) %>%
  arrange(desc(estimate)) %>%
  kable
```

```{r model_helpers}
m = function(frmla) svyglm(frmla, design=des, family=binomial(link='log'))

# explained excess risk ratio
eerr = function(m1, m2, coeff='sexfemale') {
  (1 - (exp(m2$coefficients[coeff]) - 1)/(exp(m1$coefficients[coeff]) - 1)) %>% unname
}
report_eerr = function(m1, m2) round(100 * eerr(m1, m2), 1)
```

# Nitrofurantoin

```{r nitro_models}
nitro1 = m(nitrofurantoin ~ sex)
nitro2 = m(nitrofurantoin ~ sex + uti)
```

This should be easy to explain by UTIs. They explain `r report_eerr(nitro1, nitro2)`% of the relative risk.

# Metronidazole

```{r metro_models}
metro1 = m(metronidazole ~ sex)
metro2 = m(metronidazole ~ sex + vaginitis)
```

Metronidazole consumption is increased in women. This is probably vaginitis (`61610`). So vaginitis explains `r report_eerr(metro1, metro2)`% of the relative risk. Not stupendous.

# Clindamycin

Clinda is only marginally significant in the models.

```{r clinda_models}
clinda1 = m(clindamycin ~ sex)
clinda2 = m(clindamycin ~ sex + skin)
```

Skin conditions explain `r report_eerr(clinda1, clinda2)`% of the relative risk. Also not stupendous.