### 2017-06-22 ms2/ms2.Rmd
ms2 new mtric, cleanup

---
title: "ms2"
author: "Scott Olesen"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
                      pdf.options(useDingbats=FALSE),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
import::from(broom, tidy)
import::from(tools, md5sum)
```

```{r load_data}
regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
  select(state, state_abbreviation, region)

zipcodes = read_tsv('../../db/zipcode/zipcode.tsv') %>%
  select(zipcode=zip, state)

# Use the zipcode-HRR mapping from 2011
hrr = read_tsv('../../db/hrr/hrr.tsv') %>%
  filter(year==2011) %>%
  select(zipcode, hrr) %>%
  left_join(zipcodes, by='zipcode') %>%
  left_join(regions, by='state')

years = 2011:2014
read_years = function(template_fn) {
  lapply(years, function(y) read_tsv(sprintf(template_fn, y)) %>% mutate(year=y)) %>%
    bind_rows()
}

all_ineq = read_years('data/state_ineq_all_%i.tsv') %>%
  group_by(state) %>%
  summarize(mean_cpb=mean(mean),
            sem_cpb=sd(mean)/sqrt(n()),
            mean_fnz=mean(fnz),
            sem_fnz=sd(fnz)/sqrt(n()),
            mean_nzgini=mean(nzgini),
            sem_nzgini=sd(nzgini)/sqrt(n())) %>%
  left_join(regions, by='state')

state_ineq = read_years('data/state_ineq_%i.tsv') %>%
  rename(cpb=mean, nzg=nzgini, nzr=mean_fnz_residual) %>%
  group_by(state, drug_group) %>%
  summarize_at(vars(-state, -drug_group, -total), mean) %>%
  ungroup()

hrr_ineq = read_years('data/hrr_ineq_%i.tsv') %>%
  rename(cpb=mean, nzg=nzgini, nzr=mean_fnz_residual) %>%
  group_by(hrr, drug_group) %>%
  summarize_at(vars(-hrr, -drug_group, -total), mean) %>%
  ungroup()
```

```{r load_abg, cache.extra=md5sum('bug_drug.tsv')}
resistance_groups = read_tsv('resistance_groups.tsv')
bug_names = read_tsv('bug_names.tsv')
bug_groups = read_tsv('bug_groups.tsv')
possible_bug_drug = read_tsv('bug_drug.tsv')

raw_abg = read_tsv('../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
  mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
  left_join(hrr, by=c('zipcode', 'state')) %>%
  right_join(resistance_groups, by='drug') %>%
  select(bug, raw_drug=drug, drug_group, state, hrr, percent_nonsusceptible, n_isolates)

abg = raw_abg %>%
  # shorten the bug names
  left_join(bug_names, by='bug') %>%
  select(-bug) %>% rename(bug=bug_short) %>%
  # group the different S. aureus into meta-groups depending on the drug
  left_join(bug_groups, by=c('bug', 'drug_group')) %>%
  mutate(bug_group=if_else(is.na(bug_group), bug, bug_group)) %>%
  select(-bug) %>% rename(bug=bug_group) %>%
  # keep only particular bug/drug combos
  right_join(possible_bug_drug, by=c('bug', 'drug_group'))

# get the median number of isolates for each bug/drug
median_n_isolates = abg %>%
  filter(!is.na(n_isolates)) %>%
  group_by(bug, drug_group) %>%
  summarize(median_n_isolates=as.integer(median(n_isolates))) %>%
  ungroup()

# replace the NA number of isolates with the median for that bug/drug
abg %<>%
  left_join(median_n_isolates) %>%
  mutate(n_isolates=if_else(is.na(n_isolates), median_n_isolates, n_isolates)) %>%
  select(-median_n_isolates)

# check that all the short names made it in
for (x in c('bug', 'n_isolates', 'percent_nonsusceptible')) {
  if (any(is.na(abg[[x]]))) stop(sprintf('NAs in field "%s"', x))
}

summarize_abg = function(df, place_name) {
    group_by_(df, 'bug', 'drug_group', place_name) %>%
    summarize(mean_percent_nonsusceptible=weighted.mean(percent_nonsusceptible, sqrt(n_isolates)),
              n_place_antibiograms=n()) %>%
    ungroup() %>%
    mutate(place_weight=sqrt(n_place_antibiograms))
}

state_abg = summarize_abg(abg, 'state')
hrr_abg = summarize_abg(abg, 'hrr')
```

```{r do_models}
linear_interval = function(model, level) {
  confint(model, level=level) %>%
    tidy %>%
    setNames(c('term', 'ci_low', 'ci_high'))
}

linear_model = function(df, dependent_var, explanatory_vars) {
  f = paste0(dependent_var, ' ~ ', paste0(explanatory_vars, collapse = ' + ')) %>% formula
  m = lm(formula=f, weights=place_weight, data=df)
  tidy(m) %>%
    select(term, estimate, p.value) %>%
    mutate(n=nrow(df)) %>%
    left_join(linear_interval(m, 0.95), by='term')
}

spearman_interval = function(df, rho_name, n_name, level) {
  mada::CIrho(df[[rho_name]], df[[n_name]], level=level) %>%
    tidy %>%
    setNames(c('rho', 'ci_low', 'ci_high')) %>%
    select(-rho)
}

spearman_model = function(df, dependent_var, explanatory_var) {
  cor.test(df[[dependent_var]], df[[explanatory_var]], method='spearman') %>%
    tidy %>%
    mutate(term=explanatory_var, method='spearman') %>%
    mutate(n=nrow(df)) %>%
    group_by(term, estimate, p.value, n) %>%
    do(spearman_interval(., 'estimate', 'n', 0.95)) %>%
    ungroup()
}

uni_term = 'cpb' # univariable explanatory term
multi_term = 'nzg' #'nb_prob' # multivariable explanatory term
multi_term_adjuster = 'cpb' # other multivarible explanatory term

models = function(x) {
  dep = 'mean_percent_nonsusceptible'
  bind_rows(
    linear_model(x, dep, uni_term) %>% mutate(model='uni'),
    linear_model(x, dep, c(multi_term_adjuster, multi_term)) %>% mutate(model='multi'),
    spearman_model(x, dep, uni_term) %>% mutate(model='spearman')
  ) %>%
    mutate(n_data=nrow(x))
}

results = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  group_by(bug, drug_group) %>%
  do(models(.)) %>%
  ungroup

hrr_results = left_join(hrr_abg, hrr_ineq, by=c('hrr', 'drug_group')) %>% 
  group_by(bug, drug_group) %>%
  do(models(.)) %>%
  ungroup
```

# Landscape of bug/drug correlations

## Selecting correlations to test

I join the resistance drugs into the classes shown below. (I also joined the consumption drugs into similar classes.)

`r kable(resistance_groups)`

The *S. aureus* antibiograms are listed as MSSA, MRSA, or plain old *S. aureus*. For tetracycline, Bactrim, and clindamycin, I didn't find any big difference in resistance patterns about the three groups, so I merged them all into one group for purposes of the analysis. For the cephalosporins, I joined MSSA and plain old *S. aureus*. (MRSA, on the other hand, had clear indications that the data were artificially set to 100% resistance. Intriguingly, it seemed like this was also the case for MSSA and the beta-lactams [penicillin and ampicillin], so I did not include those.)

I started by considering these `r nrow(possible_bug_drug)` bug/drug combinations on the basis of their clinical relevance and plausibility:

`r kable(possible_bug_drug)`

(In previous iterations, I filtered by the number of states/HRRs in which each bug/drug combination appeared. Weirdly, there is a relationship in the data, that better-represented bug/drug combinations are the ones with the stronger relationships [i.e., the higher $\rho$, not necessarily the lower $p$]. When using the HRR data, that kind of fancy footwork wasn't necessary, so I just didn't do it.)

## Preparing the consumption data

Unlike previous analyses, this time I treated the consumption and resistance data similarly: I averaged both across all data years. Thus, to get a state's Gini coefficient, I computed that number each year and then took the average.

## Spearman

For each bug/drug combination, I computed `mean percent nonsusceptible ~ claims per beneficiary`. A star after the bug/drug combination implies that the adjusted $p < 0.05$, and a dot means the unadjusted $p < 0.05$. Intervals show 95% confidence intervals.

```{r hrr_spearman_interval}
interval_plot = function(df) {
  arrange(df, estimate) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  mutate(sig=case_when(.$p.adj < 0.05 ~ '*',
                       .$p.value < 0.05 ~ '.',
                       TRUE ~ ''),
         bug_drug=interaction(bug, drug_group, sig),
         bug_drug=factor(bug_drug, levels=bug_drug, ordered=TRUE)) %>% 
    ggplot(aes(x=bug_drug, y=estimate, ymin=ci_low, ymax=ci_high, color=drug_group)) +
    geom_point() +
    geom_linerange() +
    coord_flip()
}

hrr_results %>%
  filter(model=='spearman') %>%
  interval_plot() +
  ylab("Spearman's rho")
```

All the demonstrably-different-from-zero relationships are positive. This jives with our general understanding that more consumption means more resistance. Note also the bunching by drug type, rather than organism.

This table has the significant and marginally significant results from the previous figure.

```{r hrr_spearman_table}
hrr_results %>%
  filter(model=='spearman') %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  filter(p.value < 0.05) %>%
  select(bug, drug_group, p.adj, p.value, rho=estimate, ci_low, ci_high) %>%
  arrange(p.value) %>%
  kable
```

Univariable linear regressions are in the Supplement. A crazy visualization of all the bug/drug combinations and resistances is also in the Supplement.

# Unevenness of consumption and antibiotic resistance

## Drugs are consumed differently in different places

```{r nzgini_cpb_drug}
cbbPalette <- c("#000000", "#D55E00", "#CC79A7", "#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")
state_ineq %>%
  ggplot(aes(x=cpb, y=nzg, color=drug_group)) +
  geom_point() +
  ylab('Gini coefficient among users') + xlab('claims per beneficiary') +
  theme_minimal() +
  scale_color_manual(values=cbbPalette)
```

## Effect of unevenness on antibiotic resistance

Here we use the same bug/drug combinations, but instead we test the multivariable model mean percent nonsusceptible ~ fraction nonzero + nonzero Gini.

There is only one significant and one marginally significant result:

```{r multi_models}
results %>%
  filter(model=='multi', term==multi_term) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  filter(p.value < 0.05) %>%
  arrange(p.value) %>%
  select(bug, drug_group, estimate, p.value, p.adj) %>%
  kable
```

```{r}
results %>%
  filter(model=='multi', term==multi_term) %>%
  interval_plot() +
  ylab('Gini coefficient')
```

Using HRRs, one of the answers remains the same, but a different marginally-significant result pops up.

```{r}
hrr_results %>%
  filter(model=='multi', term==multi_term) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  filter(p.value < 0.05) %>%
  arrange(p.value) %>%
  select(bug, drug_group, estimate, p.value, p.adj) %>%
  kable
```

```{r}
hrr_results %>%
  filter(model=='multi', term==multi_term) %>%
  interval_plot()
```

## Specific plots

Quinolone consumption predicts *E. coli* quinolone resistance:

```{r ecoli_quin}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='E. coli', drug_group=='quinolone') %>%
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  xlim(0.2, 0.45) + ylim(0, 50) +
  xlab('quinolone claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
  theme_minimal()
```

## E.coli and bactrim

```{r ecoli_bactrim}
ecoli_bactrim_dat = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') #%>%
  #filter(!is.na(mean_percent_nonsusceptible))

ecoli_bactrim_dat %>%
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
  ylim(10, 45) +
  xlab('TMP/SMX claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
  theme_minimal()
```

```{r}
ecoli_bactrim_dat %>%
  ggplot(aes(x=nzg, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  xlab('Gini coefficient)') +
  ylab('E. coli resistance') +
  theme_minimal()
```

## Pneumococcus and macrolides

```{r}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='S. pneumoniae', drug_group=='macrolide') %>%
  filter(!is.na(mean_percent_nonsusceptible)) %>% 
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  theme_minimal()
```

## E. cloacae and quinolones

```{r}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='Eb. cloacae', drug_group=='quinolone') %>%
  filter(!is.na(mean_percent_nonsusceptible)) %>%
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  theme_minimal()
```

## S. aureus and clindamycin

```{r staph_clinda}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  left_join(regions, by='state') %>%
  filter(bug=='S. aureus', drug_group=='clindamycin') %>%
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
  geom_smooth(method='lm', col='gray', se=FALSE) +
  geom_point() +
  geom_text() +
  xlab('clindamycin claims per beneficiary') + ylab('mean percent nonsusceptible S. aureus isolates') +
  theme_minimal()
```

# Supplementary results

## Amount of data vs. strength of correlation

There is a correlation between the amount of data (i.e., the number of HRRs in which there is an antibiogram for a bug/drug combination) and the correlation coefficient for that relationship. The bug/drug combinations with more data tend to be the ones with more positive $\rho$ values. This makes it advantageous to the study's power to filter the poorly-represented combinations.

```{r}
abg %>% 
  select(bug, drug_group, hrr) %>%
  distinct() %>%
  count(bug, drug_group) %>% ungroup() %>% rename(n_places=n) %>%
  right_join(filter(hrr_results, model=='spearman')) %>%
  ggplot(aes(x=n_places, y=estimate)) +
  geom_point()
```

## Metrics of unevenness

### Aggregate usage correlates with fraction of users

Error bars show the s.e.m. across the four data years.

```{r fnz_mean}
all_ineq %>%
  mutate_at(vars(matches('fnz')), function(x) x*100) %>%
  ggplot(aes(x=mean_cpb, xmin=mean_cpb-sem_cpb, xmax=mean_cpb+sem_cpb,
             y=mean_fnz, ymin=mean_fnz-sem_fnz, ymax=mean_fnz+sem_fnz,
             color=region, label=state_abbreviation)) +
  geom_point() +
  geom_linerange() +
  geom_errorbarh(height=0) +
  geom_text(nudge_x=0.04) +
  xlab('mean claims per beneficiary') +
  ylab('mean % beneficiaries with at least one claim') +
  theme_minimal()
```

```{r}
all_ineq %$%
  cor.test(mean_cpb, mean_fnz, method='spearman')
```

### The two inequality metrics not correlate well

Error bars show the s.e.m. across the four data years.

```{r nzgini_mean}
all_ineq %>%
  ggplot(aes(x=mean_fnz, xmin=mean_fnz-sem_fnz, xmax=mean_fnz+sem_fnz,
             y=mean_nzgini, ymin=mean_nzgini-sem_nzgini, ymax=mean_nzgini+sem_nzgini,
             color=region, label=state_abbreviation)) +
  geom_point() +
  geom_linerange() +
  geom_errorbarh(height=0) +
  geom_text(nudge_x=0.00) +
  xlab('mean fraction with at least one claim') +
  ylab('mean Gini coefficient') +
  theme_minimal()
```

The correlation exists but it's much less strong.

```{r}
all_ineq %$%
  cor.test(mean_fnz, mean_nzgini, method='spearman')
```

There is an intermediate correlation between claims per beneficiary and the Gini coefficient (see below), so in multivariable regressions I will use those two (rather than CPB and Gini).

```{r}
all_ineq %$% cor.test(mean_cpb, mean_nzgini, method='spearman')
```

## State-level Spearman models

```{r state_spearman_table}
results %>%
  filter(model=='spearman') %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  filter(p.value < 0.05) %>%
  arrange(p.adj) %>%
  select(bug, drug_group, estimate, p.value, p.adj, n) %>%
  kable
```

```{r state_spearman_interval}
results %>%
  filter(model=='spearman') %>%
  interval_plot() +
  ylab("Spearman's rho")
```

## Comparing Spearman models state and HRR

HRR-level Spearman correlations between consumption and resistance (almost) always underestimate the correlation reported at the state level:

```{r}
bind_rows(mutate(results, group='state'),
          mutate(hrr_results, group='hrr')) %>%
  filter(model=='spearman') %>%
  arrange(estimate) %>%
  mutate(bug_drug=interaction(bug, drug_group)) %>%
  select(bug_drug, group, estimate) %>%
  spread(group, estimate) %>%
  ggplot(aes(x=state, y=hrr, label=bug_drug)) +
  geom_point() +
  geom_text() +
  geom_abline()
```

## Univariable regressions

Like the Spearman models, the model here is `mean percent nonsusceptible ~ claims per beneficiary`, only this time it's a normal linear regression. At the HRR level:

```{r hrr_uni_table}
hrr_results %>%
  filter(model=='uni', term==uni_term) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  filter(p.value < 0.05) %>%
  arrange(p.value) %>%
  select(bug, drug_group, estimate, p.value, p.adj) %>%
  kable
```

```{r hrr_uni_interval}
hrr_results %>%
  filter(model=='uni', term==uni_term) %>%
  interval_plot()
```

And at the state level:

```{r state_uni_interval}
results %>%
  filter(model=='uni', term==uni_term) %>%
  interval_plot()
```

```{r state_uni_table}
results %>%
  filter(model=='uni', term==uni_term) %>%
  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
  filter(p.value < 0.05) %>%
  arrange(p.value) %>%
  select(bug, drug_group, estimate, p.value, p.adj, n) %>%
  kable
```

## Consumption/resistance patterns

Different bug/drug combinations have different spreads along the consumption and resistance axes, so we would expect to see different strengths of association for each one.

```{r consumption_resistance_patterns}
left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
  filter(bug %in% c('E. coli', 'Eb. cloacae', 'CoNS', 'MSSA', 'MRSA', 'S. pneumoniae')) %>%
  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, color=drug_group, shape=bug)) +
  geom_point() +
  xlab('claims per beneficiary')
```