### 2017-06-28 sex-difference/sex-difference.Rmd

---
title: "Drivers of sex differences in consumption"
author: "Scott Olesen"
date: "2/24/2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
```

# To do

- Separate beneficiaries into "high users" and "low users" based on the median number of outpatient claim dates
- Try segmenting azithromycin consumption by low/high in wo/men and by RC+/- in wo/men. We want to say that RC, more than low/high, accounts for differences in azithromycin risk, i.e., it's that women get more respiratory complaint diagnoses, not that women just have more complaints. Or vice versa.

# Background

Why do females account for 25% more antibiotic claims per capita than males? I have two easy explanations:

- More females are alive at later ages, and older people consume more antibiotics.
- Females consume more of the same antibiotics.
- Females consume different antibiotics.

# Methods

I'll use the Medicare 2011 data. Or, use all years, pool that data, and include calendar year as part of the regression.

*Outcome*: At least one claim for azithromycin, Bactrim, TMP/SMX, or nitrofurantoin. *Exposure*: A diagnosis with a UTI, COPD, or an acute RTI.

- In the first pass, look for any claim and any diagnosis.
- In a second pass, look for a diagnosis that came at most two weeks before a claim.
- Maybe, in a third pass, look for diagnosis that came at most two weeks before the first claim for that drug for the beneficiary. (Imagine someone is on year-long prophylaxis. Then a diagnosis at any time of the year will make it look like the diagnosis caused the claims.)

Before a model, I should just stratify by sex and exposure. Do we already see the signal (or lack of it)? Then we can go onto including age and calendar year as covariates.

```{r load_data}
year = 2011

pde = read_tsv(sprintf('../pde_%i.tsv', year))

sex_icd = read_tsv('sex_codes.tsv', col_types=cols(code='c'))

dx_car = sprintf('../../data/dx_car_claims_%i.tsv', year) %>%
  read_tsv(col_types=cols(diagnosis='c'))
dx_op = sprintf('../../data/dx_op_claims_%i.tsv', year) %>%
  read_tsv(col_types=cols(diagnosis='c'))

dx = bind_rows(dx_car, dx_op) %>%
  mutate(date=dmy(from_date)) %>%
  select(bene_id=BENE_ID, date, code=diagnosis) %>%
  left_join(sex_icd, by='code')

dx_counts = dx %>%
  count(code)

bene_dx = dx %>%
  select(bene_id, diagnosis_type) %>%
  distinct() %>%
  mutate(dummy=TRUE) %>%
  spread(diagnosis_type, dummy)

rc_dx = dx %>%
  filter(diagnosis_type=='acute_rc') %>%
  select(bene_id, date) %>%
  group_by(bene_id) %>%
  summarize(n_acute_rc=length(unique(date)))

total_pde = pde %>% count(bene_id) %>% rename(n_claims=n)

cc = read_feather(sprintf('../cc_%i.feather', year))

bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
  mutate(is_female=sex=='female', is_white=race=='white') %>%
  left_join(total_pde, by='bene_id') %>% replace_na(list(n_claims=0)) %>%
  left_join(bene_dx, by='bene_id') %>% replace_na(list(acute_rc=FALSE, copd=FALSE, uti=FALSE)) %>%
  left_join(select(cc, bene_id, COPD), by='bene_id') %>%
  left_join(rc_dx, by='bene_id') %>% replace_na(list(n_acute_rc=0))
```

## Diagnostic codes and counts

```{r}
sex_icd %>%
  left_join(dx_counts, by='code') %>%
  arrange(diagnosis_type, desc(n)) %>%
  mutate(n=format(n, big.mark=',')) %>%
  kable(align='lllr')
```

# Results

## Top drugs

Which drugs are the most used?

```{r top_drugs}
top_abx = pde %>%
  count(antibiotic) %>%
  arrange(desc(n)) %>%
  head(20)

kable(top_abx)
```

## Consumption of drugs by sex

Looking at top drugs by sex.

```{r drugs_by_sex}
usage_by_sex = function(drug) {
  bene %>%
    left_join(pde %>% filter(antibiotic==drug) %>% count(bene_id), by='bene_id') %>%
    replace_na(list(n=0)) %>%
    group_by(sex) %>%
    summarize(n_bene=n(),
              n_claims=sum(n),
              mean_n_claims=mean(n),
              sd_n_claims=sd(n)) %>%
    mutate(se_n_claims=sd_n_claims/sqrt(n_bene),
           drug=drug)
}

ubs = lapply(top_abx$antibiotic, usage_by_sex) %>% bind_rows
```

```{r}
denom = bene %>% count(sex) %>% rename(n_denom=n)
ubs = pde %>%
  select(bene_id, antibiotic) %>%
  distinct() %>%
  left_join(select(bene, bene_id, sex), by='bene_id') %>%
  count(sex, antibiotic) %>%
  ungroup() %>%
  rename(n_user=n) %>%
  left_join(denom, by='sex') %>%
  mutate(frac_using=n_user/n_denom) %>%
  select(antibiotic, sex, frac_using) %>%
  spread(sex, frac_using)
```

```{r drug_by_sex_plot}
ubs %>%
  mutate(highlight=antibiotic %in% c('azithromycin', 'ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'fosfomycin', 'nitrofurantoin')) %>%
  ggplot(aes(x=male, y=female, label=antibiotic, color=highlight)) +
  scale_color_manual(values=c('black', 'red')) +
  geom_abline() +
  geom_abline(slope=1.28, col='gray') +
  geom_text() +
  geom_point() +
  xlab('claims per 1,000 male beneficiaries') +
  ylab('claims per 1,000 female beneficiaries') +
  theme_minimal() +
  coord_fixed()
```

```{r drug_by_sex_plot2}
total_plot_f = function(df, highlight=character(0)) {
  df %>%
    #mutate_at(vars(total, se), mpy) %>%
    mutate(ci=se*1.96,
           ci_low=total-ci,
           ci_high=total+ci) %>%
    select(drug, sex, total, ci_low, ci_high) %>%
    gather('key', 'value', total, ci_low, ci_high) %>%
    mutate(key=case_when(.$sex=='male' & .$key=='total' ~ 'x',
                         .$sex=='male' & .$key=='ci_low' ~ 'xmin',
                         .$sex=='male' & .$key=='ci_high' ~ 'xmax',
                         .$sex=='female' & .$key=='total' ~ 'y',
                         .$sex=='female' & .$key=='ci_low' ~ 'ymin',
                         .$sex=='female' & .$key=='ci_high' ~ 'ymax')) %>%
    select(-sex) %>%
    spread(key, value) %>%
    mutate(highlight=drug %in% highlight) %>%
    ggplot(aes(x=x, xmin=xmin, xmax=xmax, y=y, ymin=ymin, ymax=ymax, label=drug, color=highlight)) +
    geom_abline(col='gray') +
    geom_point() + geom_errorbar() + geom_errorbarh() +
    scale_color_manual(values=c('black', 'red')) +
    geom_text() +
    #geom_text(nudge_x=1, nudge_y=0.5) +
    #scale_x_continuous(breaks=c(0, 5, 10), limits=c(0, 12)) +
    #scale_y_continuous(breaks=c(0, 5, 10), limits=c(0, 14)) +
    xlab('No. rx. for men (millions)') + ylab('No. rx. for women (millions)') +
    coord_fixed() +
    theme_minimal()
}

ubs %>%
  rename(total=mean_n_claims, se=se_n_claims) %>%
  total_plot_f()
```

Alternatively, I look at the mean usage (across men and women), the residual (the difference between the fraction of female beneficiaries taking the drug and male beneficiaries), and the fractional residual (the residual over the mean). I keep only those drugs with mean fraction at least 1%, and I also keep fosfomycin.

```{r}
ubs %>%
  mutate(mean_fraction=0.5*(male+female),
         residual=female-male,
         fractional_residual=residual/mean_fraction) %>%
  filter(mean_fraction > 0.01 | antibiotic=='fosfomycin') %>%
  ggplot(aes(x=mean_fraction, y=fractional_residual, label=antibiotic)) +
  geom_point() +
  geom_text()
```

This makes the same four look interesting; maybe levofloxacin and metronidazole would be next. Levo fits in with the ARTI story; metro does not.

## UTIs

Are the increased consumptions of cipro and TMP/SMX explained by UTIs? That is, do men and women who have UTIs get cipro and TMP/SMX at similar rates, and it's just that women get more UTIs than men?

### Cipro

```{r}
outcome_bene = pde %>% filter(antibiotic=='ciprofloxacin') %$% bene_id %>% unique
epi = bene %>%
  mutate(outcome=bene_id %in% outcome_bene) %>%
  count(sex, uti, outcome) %>%
  ungroup()
```

```{r, eval=FALSE}
# group_by_all_ = function(df, group_name) {
#   bind_rows(df, mutate_(df, .dots=setNames(list(~'all'), group_name))) %>%
#     group_by_(group_name)
# }
# 
# epi %>%
#   group_by_all_('sex') %>%
#   summarize(n_cases=sum(n)) %>%

```

## Azithromycin

Nitrofurantoin, ciprofloxacin, and TMP-SMX are all used as urinary anti-infectives, and UTIs more commonly affect women. This does not explain, however, the increased consumption of azithromycin.

```{r women_azithro_vs_uai, eval=FALSE}
any_in = function(lst, elts) match(elts, lst) %>% { .[!is.na(.)] } %>% { length(.) > 0 }

usage %>%
  filter(age >= 65, sex=='female') %>%
  group_by(bene_id) %>%
  mutate(used_azithro='azithromycin' %in% antibiotic,
         used_uai=any_in(antibiotic, c('nitrofurantoin', 'trimethoprim/sulfamethoxazole', 'ciprofloxacin')),
         used_both=used_azithro && used_uai,
         is_white=race=='white') %>%
  ungroup() %>%
  mutate(user_type=case_when(.$used_both ~ 'both',
                             .$used_azithro ~ 'azithro',
                             .$used_uai ~ 'UAI',
                             TRUE ~ 'none')) %>%
  group_by(bene_id) %>%
  summarize(age=age[1],
            is_white=is_white[1],
            chronic=chronic[1],
            user_type=user_type[1],
            n_claims=sum(n_claims)) %>%
  group_by(user_type) %>%
  summarize(n_beneficiaries=n(),
            mean_age=mean(age),
            percent_white=mean(is_white)*100,
            mean_n_chronic_cond=mean(chronic),
            mean_n_claims=mean(n_claims)) %>%
  arrange(desc(n_beneficiaries)) %>%
  kable
```

Does that mean that:

- women are older and therefore use more azithromycin,
- women have more chronic conditions (esp. COPD) and therefore use more, or
- women have more encounters with doctors and therefore are prescribed more?

### Chronic conditions

Cf. the chronic conditions report.

### Encounters by sex

Do women have more encounters than men? Does that explain their increased usage?

```{r load_carrier_data, eval=FALSE}
carrier = read_tsv('../../data/carrier_count_2011.tsv') %>%
  rename(bene_id=BENE_ID, n_carrier_dates=n_from_dates)

inpatient = read_tsv('../../data/inpatient_count_2011.tsv') %>%
  rename(bene_id=BENE_ID, n_inpatient_dates=n_from_dates)

outpatient = read_tsv('../../data/outpatient_count_2011.tsv') %>%
  rename(bene_id=BENE_ID, n_outpatient_dates=n_from_dates)

denom = usage %>% select(bene_id, sex, age) %>% distinct
```

```{r azithromycin_usage, eval=FALSE}
na_to_zero = function(x) if_else(is.na(x), 0L, x)

azi_usage = usage %>%
  filter(antibiotic=='azithromycin') %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims)) %>%
  left_join(denom, by='bene_id') %>%
  filter(age >= 65, sex %in% c('male', 'female')) %>%
  left_join(carrier, by='bene_id') %>%
  left_join(inpatient, by='bene_id') %>%
  left_join(outpatient, by='bene_id') %>%
  mutate(age_group=bounds_group(age, c(65, 75, 85, 95))) %>%
  mutate_at(vars(n_claims, n_carrier_dates, n_inpatient_dates, n_outpatient_dates), na_to_zero)
```

```{r azithromycin_usage_table, eval=FALSE}
azi_usage %>%
  group_by(age_group, sex) %>%
  summarize(n_beneficiaries=n(),
            mean_n_azithromycin_claims=mean(n_claims),
            med_n_carrier_dates=median(n_carrier_dates),
            med_n_outpatient_dates=median(n_outpatient_dates),
            med_n_inpatient_dates=median(n_inpatient_dates)) %>%
  kable
```

And I'll do a Poisson regression of number of azithromycin claims against age, being female, and number of carrier dates:

```{r azithromycin_usage_model, eval=FALSE}
library(broom)
azi_usage %>%
  mutate(is_female=sex=='female') %>%
  glm(n_claims ~ age + is_female + n_carrier_dates + n_inpatient_dates + n_outpatient_dates, data=., family='poisson') %>%
  tidy %>%
  mutate(pct.increase=round((exp(estimate)-1.0)*100, 2)) %>%
  kable
```

The coefficient for being female (0.05) accounts for almost the entire difference in consumption between men and women (56 per thousand). Note that age has an approximately zero correlation (i.e.,

# Supplement

## Consumption by age and sex

At every age, women consume more antibiotics than men.

```{r consumption_by_age_and_sex, eval=FALSE}
bene %>%
  left_join(total_pde, by='bene_id') %>%
  replace_na(list(n_claims=0)) %>%
  group_by(sex, age) %>%
  summarize(mean_n_claims=mean(n_claims)) %>%
  ungroup() %>%
  ggplot(aes(x=age, y=mean_n_claims, color=sex)) +
  geom_point()
```

### Sex by age

The older people are more female: age 65 is about 60% female. This stays constant around 60% (or dips slightly) until age 75, when it steadily increases up to 85-90% female at age 100. So we'll need to account for this age difference.

```{r sex_by_age, eval=FALSE}
plot_dat = usage %>%
  select(age, sex) %>%
  filter(sex %in% c('male', 'female')) %>%
  group_by(age) %>% count(sex) %>% ungroup() %>%
  spread(sex, n, fill=0L) %>%
  mutate(total=female+male, grand_total=sum(total),
         fraction_year=total/grand_total, fraction_female=female/total)

ggplot(mapping=aes(x=age, y=fraction)) +
  geom_point(data=rename(plot_dat, fraction=fraction_female)) +
  geom_line(data=mutate(plot_dat, fraction=fraction_year*10), col='red') +
  ylab('fraction female') + ylim(0, 1) +
  theme_minimal()
```

## Margins of azithromycin consumption

Maybe women take more azithromycin because usage is more concentrated in women than in men. Then we could explain the increased consumption by a small number of people taking more drug. (Yonatan points to Lady Windmere syndrome.) In fact, consumption is more even among women. As always, this is driven by the extensive fraction: more women took at least one dose:

```{r, eval=FALSE}
pde %>%
  filter(drug=='azithromycin') %>%
  count(bene_id) %>% rename(n=n_azi) %>%
  right_join(bene, by='bene_id') %>% replace_na(list(n_azi=0)) %>%
  count(sex, n_azi) %>%
  group_by(sex) %>%
  mutate(f_ppl=n/sum(n), cum_f_ppl=cumsum(f_ppl),
         cum_cons=cumsum(n_azi*n), cum_f_cons=cum_cons/max(cum_cons)) %>%
  ggplot(aes(x=cum_f_ppl, y=cum_f_cons, color=sex)) +
  geom_point() +
  geom_line() +
  xlab('Fraction of consumers') + ylab("Fraction of that sex's azithromycin consumed")
```